---
title: "How Have Shark Attacks Changed over Time?"
output: flexdashboard::flex_dashboard
runtime: shiny
---

# Loading in the Data and Packages
```{r global, include=FALSE, warning=FALSE, message=FALSE}
shark_attacks <- read.csv("Australian Shark-Incident Database Public Version.csv") #Shark Dataset.csv

# load packages in 'global' chunk so they can be shared by all users of the dashboard
library(httr2)
library(purrr)
library(tibble)
library(lubridate)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyr)
library(ggseas) #to use stat_rollapplyr

```

- first reclassify dates (1790-1850, 1850-1900, 1900-1950, 
- most common location, victim activity, site.category, distance to shore over time, proportion barchart
- 


# cleaning the data

```{r}
View(shark_attacks)
view(shark_attacks_timeline)
shark_attacks$Incident.year
view(shark_attacks$Victim.age)

shark_attacks$Victim.activity
#adding in means
shark_attacks <- shark_attacks %>%
  filter(is.na(UIN) != T) %>% #removing the empty rows
  mutate(Victim.injury = case_when(Victim.injury == "fatal" ~ "Fatal",
                                   Victim.injury == "injured" ~ "Injured",
                                   Victim.injury == "Injured" ~ "Injured",
                                   Victim.injury == "uninjured" ~ "Uninjured")) %>%
  mutate(Provoked.unprovoked = case_when(Provoked.unprovoked == "provoked" ~ "Provoked",
                                   Provoked.unprovoked == "unprovoked" ~ "Unprovoked")) %>%
  mutate(Victim.gender = case_when(Victim.gender == "female" ~ "Female",
                                   Victim.gender == "male" ~ "Male")) %>%
  mutate(Injury.severity = case_when(Injury.severity == "abrasion" ~ "Abrasion", Injury.severity == "major lacerations" ~ "Major Lacerations", Injury.severity == "minor lacerations" ~ "Minor Lacerations", Injury.severity == "other: teeth marks" ~ "Other: Teeth Marks", Injury.severity == "punctures" ~ "Punctures", Injury.severity == "surface wound" ~ "Surface Wound")) %>%
  mutate(Site.category = case_when(Site.category == "coastal" ~ "Coastal",
                                   Site.category == "Coastal" ~ "Coastal",
                                   Site.category == "estuary/harbour" ~ "Estuary/Harbour",
                                   Site.category == "island open ocean" ~ "Island Open Ocean", 
                                   Site.category == "ocean/pelagic" ~ "Ocean/Pelagic",
                                   Site.category == "other: fish farm" ~ "Other: Fish Farm",
                                   Site.category == "river" ~ "River")) %>%
  filter(Incident.year != "1791") %>%
  mutate(Victim.activity = case_when(Victim.activity == "boarding" ~ "Boarding",
                                   Victim.activity == "diving" ~ "Diving",
                                   Victim.activity == "fishing" ~ "Fishing",
                                   Victim.activity == "motorised boating" ~ "Motorised Boating",
                                   Victim.activity == "other: standing in water" ~ "Standing in Water",
                                   Victim.activity == "other:floating" ~ "Floating",
                                   Victim.activity == "snorkeling" ~ "Snorkeling", 
                                   Victim.activity == "spearfishing" ~ "Spearfishing",
                                   Victim.activity == "swimming" ~ "Swimming",
                                   Victim.activity == "unmotorised boating" ~ "Unmotorised Boating")) %>%
  filter(Incident.year != "1791") %>%
  group_by(Incident.year) %>%
   #creating mean columns
  mutate(Shark.length.m.mean = mean(Shark.length.m)) %>%
  mutate(No.sharks.mean = mean(No.sharks)) %>%
  mutate(Victim.age.mean = mean(Victim.age)) %>%
  #creating date groupings
  mutate(Year = case_when((Incident.year > 1800 & Incident.year <= 1850) ~ "1800-1850", 
                          (Incident.year > 1850 & Incident.year <= 1900) ~ "1850-1900",
                          (Incident.year > 1900 & Incident.year <= 1950) ~ "1900-1950",
                          (Incident.year > 1950 & Incident.year <= 2000) ~ "1950-2000",
                          (Incident.year > 2000 & Incident.year <= 2022) ~ "2000-2022")) 

#counting the number of sharks every 50 years
shark_attacks <- shark_attacks %>%
  group_by(Year) %>%
  mutate(Number_Attacks = n())
  
#creating a dataframe for timeline data
shark_attacks_timeline <- shark_attacks %>%
  mutate(Victim.age = as.ts(Victim.age)) %>% 
  mutate(Incident.month = as.character(Incident.month)) %>% 
  mutate(Incident.year = as.character(Incident.year)) %>%
 # mutate(Incident.month_year = make_date(year = Incident.year, month=Incident.month)) 
  unite("Incident.month_year", Incident.month, Incident.year, sep = "/", remove = F) %>% #remove = F, na.rm = T
  mutate(Incident.month_year = paste0("01/", Incident.month_year),
         Incident.month_year = as.Date(Incident.month_year, format = "%d/%m/%Y"))
  

```

# Timeline

```{r}
class(shark_attacks_timeline$Shark.length.m)
shark_attacks_timeline$Incident.month_year
#start and stop times

start_time <- as.Date("1900-01-01")
end_time <- as.Date("2022-03-01")

start.end <- c(start_time,end_time)
start.end


#age 
ggplot(data = shark_attacks_timeline, aes(x = Incident.month_year, y = Victim.age)) +
  geom_line()+ 
  labs(x = "Date", y = "Victim Age") +
  scale_x_date(date_breaks = "10 years", date_labels = "%m/%Y", limits = start.end) + 
  stat_rollapplyr(width = 50, align = "right", linewidth = 1, color = "red") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8))

#shark length
ggplot(data = shark_attacks_timeline, aes(x = Incident.month_year, y = Shark.length.m)) +
  geom_line()+ 
  labs(x = "Date", y = "Shark Length (m)") +
  scale_x_date(date_breaks = "25 years", date_labels = "%m/%Y", limits = start.end) + 
  stat_rollapplyr(width = 50, align = "right", linewidth = 1, color = "red") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8))

#number of sharks during the attacks
ggplot(data = shark_attacks_timeline, aes(x = Incident.month_year, y = No.sharks)) +
  geom_line()+ 
  labs(x = "Date", y = "Number of Sharks") +
  scale_x_date(date_breaks = "25 years", date_labels = "%m/%Y", limits = start.end) + 
  stat_rollapplyr(width = 50, align = "right", linewidth = 1, color = "red") +
  ylim(c(0,10)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8))


```


- figure out the bet way to adjust the x limits (is it by having the user change the start and end values?)
- things user could change
  - length of time viewed
  - y variable
  - moving average variable

# barcharts graph -- Calculaiting the proportions and means

```{r}

#making the means
View(shark_attacks)

shark_attacks_proportions_means <- shark_attacks %>%
  group_by(Year) %>%  #  group by the mfr variable
  summarise(mean_victim_age = mean(Victim.age),
            count_victim_age = n()) %>% 
  mutate(total_fat = sum(count_fat),  #  compute total number of observations
         std_error_fat = sqrt(mean_fat * ((1 - mean_fat) / total_fat)),  #  compute standard error of each proportion
         lower_fat = mean_fat - 1.96 * std_error_fat,  #  compute CI lower bound
         upper_fat = mean_fat + 1.96 * std_error_fat) %>%  #  compute CI upper bound
  mutate(total_inj = sum(count_inj),  #  compute total number of observations
         std_error_inj = sqrt(mean_inj * ((1 - mean_inj) / total_inj)),  #  compute standard error of each proportion
         lower_inj = mean_inj - 1.96 * std_error_inj,  #  compute CI lower bound
         upper_inj = mean_inj + 1.96 * std_error_inj)  #  compute CI upper bound

shark_attacks$Victim.age
#making proportion
#  Add the following information to the cereal dataset:
#    Proportions and percentages corresponding to each category
#    The standard error on the proportions or percentages corresponding to each
#      category
#    Lower bound of an (approximate) 95% confidence interval around the true 
#      proportion in each category
#    Upper bound of an (approximate) 95% confidence interval around the true 
#      proportion in each category

#- Victim.injury, Site.category, Provoked.unprovoked, Victim.activity, Victim.gender, Injury.severity
view(shark_attacks_proportions)

shark_attacks_proportions<- shark_attacks %>%  #  Start with the cereal data.frame
  group_by(Year) %>%  #  group by the year variable
  summarise(count_Victim.injury = n(),
            count_Site.category = n(),
            count_Provoked.unprovoked = n(),
            count_Victim.activity = n(),
            count_Victim.gender = n(),
            count_Injury.severity = n()) %>%  #  summarize by calculating count of each manufacturer
  
  mutate(total_Victim.injury = sum(count_Victim.injury),  #  compute total number of observations
         proportion_Victim.injury = count_Victim.injury / total_Victim.injury,  #  compute proportions
         std_error_Victim.injury = sqrt(proportion_Victim.injury * (1-proportion_Victim.injury) / total_Victim.injury),  #  compute standard error of each proportion
         lower_Victim.injury = proportion_Victim.injury - 1.96 * std_error_Victim.injury,  #  compute CI lower bound
         upper_Victim.injury = proportion_Victim.injury + 1.96 * std_error_Victim.injury) %>%
  
  mutate(total_Site.category = sum(count_Site.category),  
         proportion_Site.category = count_Site.category / total_Site.category,  
         std_error_Site.category = sqrt(proportion_Site.category * (1-proportion_Site.category) / total_Site.category),  
         lower_Site.category = proportion_Site.category - 1.96 * std_error_Site.category,  
         upper_Site.category = proportion_Site.category + 1.96 * std_error_Site.category) %>%
  
  mutate(total_Provoked.unprovoked = sum(count_Provoked.unprovoked),  
         proportion_Provoked.unprovoked = count_Provoked.unprovoked / total_Provoked.unprovoked, 
         std_error_Provoked.unprovoked = sqrt(proportion_Provoked.unprovoked * (1-proportion_Provoked.unprovoked) / proportion_Provoked.unprovoked),  
         lower_Provoked.unprovoked = proportion_Provoked.unprovoked - 1.96 * std_error_Provoked.unprovoked,  
         upper_Provoked.unprovoked = proportion_Provoked.unprovoked + 1.96 * std_error_Provoked.unprovoked) %>%
  
  mutate(total_Victim.activity = sum(count_Victim.activity),  
         proportion_Victim.activity = count_Victim.activity / total_Victim.activity,  
         std_error_Victim.activity = sqrt(proportion_Victim.activity * (1-proportion_Victim.activity) / total_Victim.activity),  
         lower_Victim.activity = proportion_Victim.activity - 1.96 * std_error_Victim.activity,  
         upper_Victim.activity = proportion_Victim.activity + 1.96 * std_error_Victim.activity) %>%
  
  
   mutate(total_Victim.gender = sum(count_Victim.gender),  
         proportion_Victim.gender = count_Victim.gender / total_Victim.gender,  
         std_error_Victim.gender = sqrt(proportion_Victim.gender * (1-proportion_Victim.gender) / total_Victim.gender),  
         lower_Victim.gender = proportion_Victim.gender - 1.96 * std_error_Victim.gender,  
         upper_Victim.gender = proportion_Victim.gender + 1.96 * std_error_Victim.gender) %>%
  
  
  mutate(total_Injury.severity = sum(count_Injury.severity),  
         proportion_Injury.severity = count_Injury.severity / total_Injury.severity,  
         std_error_Injury.severity = sqrt(proportion_Injury.severity * (1-proportion_Injury.severity) / total_Injury.severity),  
         lower_Injury.severity = proportion_Injury.severity - 1.96 * std_error_Injury.severity,  
         upper_Injury.severity = proportion_Injury.severity + 1.96 * std_error_Injury.severity) 
  
```

# computing proportions individually

```{r}
#individually

#Victim.injury
shark_attacks_proportions_Victim.injury <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.injury != "") %>%
  group_by(Year, Victim.injury) %>%  #  group by the year variable
  summarise(count_Victim.injury = n()) %>% 
  mutate(total_Victim.injury = sum(count_Victim.injury),  #  compute total number of observations
         proportion_Victim.injury = count_Victim.injury / total_Victim.injury,  #  compute proportions
         std_error_Victim.injury = sqrt(proportion_Victim.injury * (1-proportion_Victim.injury) / total_Victim.injury),  #  compute standard error of each proportion
         lower_Victim.injury = proportion_Victim.injury - 1.96 * std_error_Victim.injury,  #  compute CI lower bound
         upper_Victim.injury = proportion_Victim.injury + 1.96 * std_error_Victim.injury) 
  

#Site.category
shark_attacks_proportions_Site.category <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Site.category != "") %>%
  group_by(Year, Site.category) %>%  #  group by the year variable
  summarise(count_Site.category = n()) %>%  
  mutate(total_Site.category = sum(count_Site.category),  
         proportion_Site.category = count_Site.category / total_Site.category,  
         std_error_Site.category = sqrt(proportion_Site.category * (1-proportion_Site.category) / total_Site.category),  
         lower_Site.category = proportion_Site.category - 1.96 * std_error_Site.category,  
         upper_Site.category = proportion_Site.category + 1.96 * std_error_Site.category)
  

#Provoked.unprovoked
shark_attacks_proportions_Provoked.unprovoked <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Provoked.unprovoked != "") %>%
  group_by(Year, Provoked.unprovoked) %>%  #  group by the year variable
  summarise(count_Provoked.unprovoked = n()) %>%  
  mutate(total_Provoked.unprovoked = sum(count_Provoked.unprovoked),  
         proportion_Provoked.unprovoked = count_Provoked.unprovoked / total_Provoked.unprovoked, 
         std_error_Provoked.unprovoked = sqrt(proportion_Provoked.unprovoked * (1-proportion_Provoked.unprovoked) / total_Provoked.unprovoked),  
         lower_Provoked.unprovoked = proportion_Provoked.unprovoked - 1.96 * std_error_Provoked.unprovoked,  
         upper_Provoked.unprovoked = proportion_Provoked.unprovoked + 1.96 * std_error_Provoked.unprovoked)
  

#Victim.activity
shark_attacks_proportions_Victim.activity <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.activity != "") %>%
  group_by(Year, Victim.activity) %>%  #  group by the year variable
  summarise(count_Victim.activity = n()) %>%  
  mutate(total_Victim.activity = sum(count_Victim.activity),  
         proportion_Victim.activity = count_Victim.activity / total_Victim.activity,  
         std_error_Victim.activity = sqrt(proportion_Victim.activity * (1-proportion_Victim.activity) / total_Victim.activity),  
         lower_Victim.activity = proportion_Victim.activity - 1.96 * std_error_Victim.activity,  
         upper_Victim.activity = proportion_Victim.activity + 1.96 * std_error_Victim.activity)


#Victim.gender 
shark_attacks_proportions_Victim.gender <- shark_attacks %>% #  Start with the data.frame
  filter(Victim.gender != "") %>%
  group_by(Year, Victim.gender) %>%  #  group by the year variable
  summarise(count_Victim.gender = n()) %>%    
  mutate(total_Victim.gender = sum(count_Victim.gender),  
         proportion_Victim.gender = count_Victim.gender / total_Victim.gender,  
         std_error_Victim.gender = sqrt(proportion_Victim.gender * (1-proportion_Victim.gender) / total_Victim.gender), 
         lower_Victim.gender = proportion_Victim.gender - 1.96 * std_error_Victim.gender,  
         upper_Victim.gender = proportion_Victim.gender + 1.96 * std_error_Victim.gender)
  
  
#Injury.severity
shark_attacks_proportions_Injury.severity <- shark_attacks %>%  #Start with the cereal data.frame
  filter(Injury.severity != "") %>%
  group_by(Year, Injury.severity) %>%  #  group by the year variable
  summarise(count_Injury.severity = n()) %>%  
  mutate(total_Injury.severity = sum(count_Injury.severity),  
         proportion_Injury.severity = count_Injury.severity / total_Injury.severity,  
         std_error_Injury.severity = sqrt(proportion_Injury.severity * (1-proportion_Injury.severity) / total_Injury.severity), 
         lower_Injury.severity = proportion_Injury.severity - 1.96 * std_error_Injury.severity,  
         upper_Injury.severity = proportion_Injury.severity + 1.96 * std_error_Injury.severity) 
  




```

# proportional graph -- Making the graphs

```{r}
#- Victim.injury, Site.category, Provoked.unprovoked, Victim.activity, Victim.gender, Injury.severity

#number of attackes
ggplot(data = shark_attacks, aes(x = Year, y = Number_Attacks)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency of Shark Attacks")

 #Victim.injury

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Victim.injury)) +
  geom_bar(stat = "identity") +
  labs(y = "Proportion of Victim Injury Type by 50 Years")
  geom_errorbar(aes(ymin = lower_Victim.injury, ymax = upper_Victim.injury), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Victim.injury, aes(x = Year, y = count_Victim.injury, fill = Victim.injury)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "Victim Injury Type", 
       y = "Frequency of Victim Injuries")


#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Victim.injury, aes(x = Year, y = proportion_Victim.injury, fill = Victim.injury)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Site Category",
       title = "Proportion of Shark Attack Victim Injury Types Across Time") +
  geom_errorbar(aes(ymin = lower_Victim.injury, ymax = upper_Victim.injury), color = "red", position = "dodge")

#Site.category

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Site.category)) +
  geom_bar(stat = "identity") +
  labs(Y = "Proportion of Site Category")+
  geom_errorbar(aes(ymin = lower_Site.category, ymax = upper_Site.category), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Site.category, aes(x = Year, y = count_Site.category, fill = Site.category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency")

#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Site.category, aes(x = Year, y = proportion_Site.category, fill = Site.category)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Site Category",
       title = "Proportion of Site Category of Shark Attacks Across Time") +
  geom_errorbar(aes(ymin = lower_Site.category, ymax = upper_Site.category), color = "red", position = "dodge")

#Provoked.unprovoked

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Provoked.unprovoked)) +
  geom_bar(stat = "identity") +
  labs(Y = "Proportion of Provoked vs. Unprovoked Shark Attacks")
  geom_errorbar(aes(ymin = lower_Provoked.unprovoked, ymax = upper_Provoked.unprovoked), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Provoked.unprovoked, aes(x = Year, y = count_Provoked.unprovoked, fill = Provoked.unprovoked)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency")

#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Provoked.unprovoked, aes(x = Year, y = proportion_Provoked.unprovoked, fill = Provoked.unprovoked)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Provoked?",
       title = "Proportion of Provoked vs. Unprovoked Shark Attacks Across Time") +
  geom_errorbar(aes(ymin = lower_Provoked.unprovoked, ymax = upper_Provoked.unprovoked), color = "red", position = "dodge")

#Victim.activity

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Victim.activity)) +
  geom_bar(stat = "identity") +
  labs(Y = "Proprortion of Victim Activity")
  geom_errorbar(aes(ymin = lower_Victim.activity, ymax = upper_Victim.activity), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Victim.activity, aes(x = Year, y = count_Victim.activity, fill = Victim.activity)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency")

#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Victim.activity, aes(x = Year, y = proportion_Victim.activity, fill = Victim.activity)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Victim Activity",
       title = "Proportion of Shark Attack Victim Activities Across Time") +
  geom_errorbar(aes(ymin = lower_Victim.activity, ymax = upper_Victim.activity), color = "red", position = "dodge")


#Victim.gender

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Victim.gender)) +
  geom_bar(stat = "identity") +
  labs(Y = "Proprortion of Victim Gender")
  geom_errorbar(aes(ymin = lower_Victim.gender, ymax = upper_Victim.gender), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Victim.gender, aes(x = Year, y = count_Victim.gender, fill = Victim.gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency")

#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Victim.gender, aes(x = Year, y = proportion_Victim.gender, fill = Victim.gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Victim Gender",
       title = "Proportion of Shark Attack Victim Genders Across Time") +
  geom_errorbar(aes(ymin = lower_Victim.gender, ymax = upper_Victim.gender), color = "red", position = "dodge")

#Injury.severity

#basic proportion
ggplot(data = shark_attacks_proportions, aes(x = Year, y = proportion_Injury.severity)) +
  geom_bar(stat = "identity") +
  labs(Y = "Proprortion of Injury Severity")
  geom_errorbar(aes(ymin = lower_Injury.severity, ymax = upper_Injury.severity), color = "red")
  
#counts side by side barchart
ggplot(data = shark_attacks_proportions_Injury.severity, aes(x = Year, y = count_Injury.severity, fill = Injury.severity)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Frequency")

#Proportion side by side barchart
ggplot(data = shark_attacks_proportions_Injury.severity, aes(x = Year, y = proportion_Injury.severity, fill = Injury.severity)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion",
       fill = "Injury Severity",
       title = "Proportion of Shark Attack Injury Severities Across Time") +
  geom_errorbar(aes(ymin = lower_Injury.severity, ymax = upper_Injury.severity), color = "red", position = "dodge")




```
