---
title: "Shark Attacks in Australia"
date: "12/11/2024"
authors: "Erica Loomis (esloom25), Timothy Pratt (tjprat25), and Rebecca Wanger (rswang25)"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: 
      version: 4
      bg: "#101010"
      fg: "#8db8fc" 
      primary: "#8db8fc"
      base_font: 
        DIN Alternate
runtime: shiny
---

```{r}
## Connecting to the Shiny App Account
# rsconnect::setAccountInfo(name='ter1',
# 			  token='0163EF6118B243DE8D0042631C039930',
# 			  secret='<Secret>')

## Actually running and deploying the app

library(shiny)
library(ggplot2)
#runApp()

#deploying the app

library(rsconnect)
#deployApp()

#link: https://ter1.shinyapps.io/sc326-tre-interactive-graphics-project/
```

```{r global, include=FALSE, warning=FALSE, message=FALSE}
# loading in data in 'global' chunk so it can be shared by all users of the dashboard
shark_attacks <- read.csv("Shark Dataset.csv")

# load packages in 'global' chunk so they can be shared by all users of the dashboard
library(httr2)
library(purrr)
library(tibble)
library(lubridate)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyr)
library(ggseas) #to use stat_rollapplyr
library(plotly)
library(forcats)

interactive_theme <- theme_gray() +# White background, black and white theme
  theme(plot.title = element_text(family = "Georgia", face = "bold", color = "#0a144a"),
        axis.title = element_text(family = "DIN Alternate", color = "#0a144a"),
        axis.text = element_text(color = "#0a144a"),
        legend.title = element_text(family = "DIN Alternate", face = "bold",  color = "#0a144a"),
        legend.text = element_text(family = "DIN Alternate", color = "#0a144a"),
        legend.box.margin = margin(6, 6, 6, 6),
        legend.background = element_rect(fill = "#d1d4e3"),
        panel.background = element_rect(fill = "#d1d4e3"),
        plot.background = element_rect(fill = "#d1d4e3"))
```

Introduction
=====================================
```{r}
fluidPage(
  titlePanel("Australian Shark Attacks & Safety: An Interactive Analysis"),
  
  fluidRow(
    column(12,
      h3("Welcome to Australia!"),
      p("Shark attacks, while rare, are a significant cause for concern for beachgoers, swimmers, and surfers along Australia's coast. In this application, we analyze various factors contributing to shark attacks in Australia, including locations, timing, and shark species. Our goal is to help visitors and locals alike understand the patterns behind shark attacks and make informed decisions about water safety."),
      p("Through this interactive dashboard, you'll be able to explore historical data on shark attacks, view trends, and identify activities where safety measures may be most needed. Whether you're a tourist or a local, this tool aims to promote awareness and enhance safety practices when enjoying Australia's beautiful coastlines."),
      p("We hope this resource helps you have a safer and more informed experience in Australia!"),
      p("Click the buttons along the top bar to start exploring different graphs.")
    )
  ),
  
  fluidRow(
    column(6,
      h4("Data Source"),
      p("The data used in this application is sourced from the publicly available Australian Shark-Incident Database (ASID) published by the Taronga Conservation Society Australia."),
      p("The full dataset can be accessed here: https://taronga.org.au/conservation-and-science/australian-shark-incident-database#thedatabase")
    ),
    column(6,
      h4("Authors"),
      p("This application was developed by:"),
      tags$ul(
        tags$li("Erica Loomis"),
        tags$li("Timothy Pratt"),
        tags$li("Rebecca Wanger"),
      )
    )
  )
)


```

How fatal is this shark?
=====================================  

```{r global_species, include=FALSE, warning=FALSE, message=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(tidyverse)
library(flexdashboard)
library(plotly)
library(stringr)
library(knitr)

shark <- read.csv("Shark Dataset.csv")

```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("shark_species_EL", label = "Attacks based on shark species", 
            choices =  c("All" = "all",
                         "White shark" = "white shark",
                         "Tiger shark" = "tiger shark",
                         "Bull shark" = "bull shark",
                         "Whaler shark" = "whaler shark",
                         "Wobbegong" = "wobbegong",
                         "Broadnose sevengill shark" = "broadnose sevengill shark",
                         "Dusky shark" = "dusky shark",
                         "Bronze whaler shark" = "bronze whaler shark",
                         "Dogfish" = "dogfish",
                         "Shortfin mako shark" = "shortfin mako shark",
                         "School shark" = "school shark",
                         "Grey reef shark" = "grey reef shark",
                         "Whitetip reef shark" = "whitetip reef shark",
                         "Hammerhead shark" = "hammerhead shark ",
                         "Grey nurse shark" = "grey nurse shark",
                         "Sevengill shark" = "sevengill shark",
                         "Galapagos shark" = "galapagos shark",
                         "Silvertip shark" = "silvertip shark",
                         "Blind shark" = "blind shark",
                         "Blacktip reef shark" = "blacktip reef shark",
                         "Port jackson shark" = "port jackson shark",
                         "Lemon shark" = "lemon shark"))

selectInput("provoke_EL", label = "Attacks if provoked or unprovoked", 
            choices = c("Provoked and Unprovoked" = "both",
                        "Provoked" = "provoked",
                        "Unprovoked" = "unprovoked"))
```

Data source: Australian Shark-Incident Database

```{r}

renderImage({
    if(input$shark_species_EL == "all"){
      list(
          src = file.path(paste("shark.jpg")),
          contentType = "image/jpeg",
          width = 230,
          alt = "Picture of a shark"
        )
    } else {
        list(
          src = file.path(paste(input$shark_species_EL, ".jpg")),
          contentType = "image/jpeg",
          width = 230,
          alt = "Picture of a shark"
        )
    }
  }, deleteFile = FALSE)


```

Row
-----------------------------------------------------------------------
### Percent Fatal

```{r}
fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup()

renderGauge({
  
if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      mutate(total_attacks = sum(sum_species),
             percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))

  } else { # and they either provoke or unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      mutate(total_attacks = sum(sum_species),
           percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } 
} else { # if is is a specific shark species
  if(input$provoke_EL == "both") { #and they provoke and unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species),
             percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } else { # and they either provoke or unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species),
           percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
        group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } 
}
  rate <- fatal_percent_gauge$percent_fatal # assign value 
  gauge(rate, min = 0, max = 100, symbol = '%', gaugeSectors( # render the gauge
    success = c(0, 39), warning = c(40, 79), danger = c(80, 100)
  ))
})
```

### Number of Attacks in this Situation
```{r}
fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup()

renderValueBox({
  
  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
}
  attacks <- as.character(fatal_percent_value$total_attacks) #assign value
  valueBox(attacks, # render the valuebox
           icon = "ion-alert-circled",
           color = "#0a144a")
})
```

### Number of Deaths
```{r}

fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup() %>% 
  filter( Shark.common.name != "") 

renderValueBox({
  
  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
}
  deaths <- fatal_percent_value$total_attacks #assign value
  valueBox(deaths, # render the valuebox
           icon = "ion-alert-circled",
           color = "#67000d")
})
```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### All species

```{r}
  shark_bubble <- shark %>%
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>%
  summarize(attack_species_provoke_fatal = n()) %>%
  mutate(species_provoke = paste(Shark.common.name, Provoked.unprovoked, sep=" ")) %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  mutate(attack_species_provoke = sum(attack_species_provoke_fatal)) %>%
  group_by(Shark.common.name, fatal) %>%
  mutate(attack_species_fatal = sum(attack_species_provoke_fatal)) %>%
  group_by(Shark.common.name) %>%
  mutate(attack_species = sum(attack_species_provoke_fatal), # all attacks by species
         all_percent = ifelse(fatal == "fatal", attack_species_fatal / attack_species, 0), # calculating the percent fatal for all
         provoke_percent = ifelse(fatal == "fatal", attack_species_provoke_fatal / attack_species_provoke, 0)) %>%
  filter( Shark.common.name != "") %>%
  mutate(Shark.common.name = str_to_title(Shark.common.name))


renderPlotly({ #going into the if statements

  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    
    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent)) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point(size = 3) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } else { # and they either provoke or unprovoke
    
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent)) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point(size = 3) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = paste("Which Shark Species Attack is Most Common and Fatal?-", input$provoke_EL)) +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke

    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent)) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
  
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } else { # and they either provoke or unprovoke
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent)) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } 
}


})
```

### Top 4 species

```{r}

renderPlotly({ #going into the if statements

  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    
    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent)) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%")) %>% 
        filter(Shark.common.name == "White Shark"| 
                 Shark.common.name == "Tiger Shark" | 
                 Shark.common.name == "Wobbegong" | 
                 Shark.common.name == "Bull Shark")
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), size = attack_species, color = percent, text = text)) +
  geom_point() +
  scale_size(range = c(1, 4), name="Number of Attacks)") +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } else { # and they either provoke or unprovoke
    
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent)) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%")) %>%
      filter(Shark.common.name == "White Shark"| 
               Shark.common.name == "Tiger Shark" | 
               Shark.common.name == "Wobbegong" | 
               Shark.common.name == "Bull Shark")
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), size = attack_species, color = percent, text = text)) +
  geom_point() +
  scale_size(range = c(1, 10), name="Number of Attacks)") +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = paste("Which Shark Species Attack is Most Common and Fatal?-", input$provoke_EL)) +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke

    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent)) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%")) %>% 
      filter(Shark.common.name == "White Shark"| 
               Shark.common.name == "Tiger Shark" | 
               Shark.common.name == "Wobbegong" | 
               Shark.common.name == "Bull Shark")
  
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } else { # and they either provoke or unprovoke
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent)) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%")) %>% 
      filter(Shark.common.name == "White Shark"| 
               Shark.common.name == "Tiger Shark" | 
               Shark.common.name == "Wobbegong" | 
               Shark.common.name == "Bull Shark")
    
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } 
}


})
```

### Shark Length
```{r}
renderPlot({ #going into the if statements

  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    
    fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal"))
    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
  scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
    
  } else { # and they either provoke or unprovoke
    
  fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
    filter(Provoked.unprovoked == input$provoke_EL) # only taking the provoked or unprovoked options

  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke

  fatal_shark <- shark %>% 
    mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
      filter(Shark.common.name == input$shark_species_EL)


    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  
  } else { # and they either provoke or unprovoke
    
fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
    filter(Provoked.unprovoked == input$provoke_EL) %>%  # only taking the provoked or unprovoked options
    filter(Shark.common.name == input$shark_species_EL)


    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  } 
}


})
```

Which body part are you most likely to injure?
=====================================  

```{r global_body, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
library(stringr)
library(devtools)
install_github("josedv82/bdgramR")
library(bdgramR)

shark <- read.csv("Shark Dataset.csv")

```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("gender_EL", label = "Select a gender", 
            choices =  c("All" = "all",
                         "Male" = "male",
                         "Female" = "female"))

sliderInput("age_EL", label = "Select your age",
            min = 0, max = 75, value = 30, step = 1)

radioButtons("bar_EL", "Bar Chart type",
               c("Bar chart of all age ranges" = "all",
                 "Bar chart of your age range" = "specific"))


```

Data source: Australian Shark-Incident Database

Row {data-height=100}
-----------------------------------------------------------------------

### Looks like...
Watch out for your legs! They seem to be the most common part of the body injured by people.

Row 
-----------------------------------------------------------------------
### Body part injuries by age chunk

```{r}
renderPlot({
  
body_injury <- shark %>%
  mutate(body_1 = str_split_fixed(Injury.location, ", ", 4)[ ,1],
         body_2 = str_split_fixed(Injury.location, ", ", 4)[ ,2],
         body_3 = str_split_fixed(Injury.location, ", ", 4)[ ,3],
         body_4 = str_split_fixed(Injury.location, ", ", 4)[ ,4],
         age_range = case_when(Victim.age <= 10 ~ "0-10", 
                              (10 < Victim.age) & (Victim.age <= 20)  ~ "11-20",
                              (20 < Victim.age) & (Victim.age <= 30)  ~ "21-30",
                              (30 < Victim.age) & (Victim.age <= 40)  ~ "31-40",
                              (40 < Victim.age) & (Victim.age <= 50)  ~ "41-50",
                              (50 < Victim.age) & (Victim.age <= 60)  ~ "51-60",
                              (60 < Victim.age) & (Victim.age <= 70)  ~ "61-70",
                              (70 < Victim.age) & (Victim.age <= 75)  ~ "71-75",
                              Victim.age == " "                ~ NA)) 
if (input$gender_EL == "all") { #if it's all genders
  
} else { #if it's a specific gender
  body_injury <- body_injury %>% 
    filter(Victim.gender == input$gender_EL)
}

body_1 <- body_injury %>% 
  group_by(body_1, age_range) %>% 
  rename(body = body_1) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_2 <- body_injury %>% 
  group_by(body_2, age_range) %>% 
  rename(body = body_2) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_3 <- body_injury %>% 
  group_by(body_3, age_range) %>% 
  rename(body = body_3) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_4 <- body_injury %>% 
  group_by(body_4, age_range) %>% 
  rename(body = body_4) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_12 <- rbind(body_1, body_2)
body_123 <- rbind(body_12, body_3)
body_1234 <- rbind(body_123, body_4)

body_total <- body_1234 %>% 
  group_by(body, age_range) %>%
  mutate(body = ifelse(body == "left leg", "leg", body),
         body = ifelse(body == "left foot", "foot", body)) %>% 
  summarize(Number_injury = sum(sum)) %>% 
  mutate(overall = case_when(body == "ankle" ~ "foot",
                            body == "calf" ~ "leg",
                            body == "hand" ~ "arm",
                            body == "leg" ~ "leg",
                            body == "other: half of body recovered" ~ "",
                            body == "other: uninjured" ~ "",
                            body == "right" ~ "",
                            body == "shoulder " ~ "shoulder",
                            body == "torso" ~ "torso",
                            body == "arm" ~ "arm",
                            body == "foot" ~ "foot",
                            body == "head" ~ "head",
                            body == "other: body not recovered" ~ "",
                            body == "other: multiple bites" ~ "",
                            body == "pelvic region" ~ "pelvic region",
                            body == "shoulder" ~ "shoulder",
                            body == "thigh" ~ "leg"),
         body = ifelse(grepl("other", body, fixed = TRUE), "", body)) %>% 
  filter(overall != "" ) %>% 
  na.omit()

final_body_bar <- body_total %>% 
    group_by(age_range) %>% 
  filter(Number_injury == max(Number_injury))

final_body_bar$age_range <- factor(final_body_bar$age_range, levels = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-75"))

body_bar <- ggplot(final_body_bar, aes(x = age_range, y = Number_injury, fill = body)) +
  geom_col() +
  labs(x = "Age Range",
       y = "Number of Injuries for this Body Part",
       title = "Which Body Part was Most Injured by Age Ranges?",
       fill = "Body Part Injured") +
  scale_fill_manual(values = c("#44AA99", "#CC6677", "#493C86", "#DDCC77"))

if (input$bar_EL == "all") { #if it's all barchart
    
 specific_age <- final_body_bar %>% 
  filter(age_range == case_when(input$age_EL <= 10 ~ "0-10", 
                              (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                              (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                              (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                              (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                              (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                              (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                              (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75"))

body_bar + geom_col(data = specific_age, color = "#a50f15", size = 1.5) + 
  guides(color = "none") +
  interactive_theme
  
} else { #if it's a specific age barchart

body_total_age <- body_total %>% 
  group_by(body) %>% 
  filter(age_range == case_when(input$age_EL <= 10 ~ "0-10", 
                               (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                               (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                               (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                               (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                               (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                               (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                               (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75"))

body_bar <- ggplot(body_total_age, aes(x = reorder(body, Number_injury), y = Number_injury)) +
  geom_col(fill = "#a50f15") +
  labs(x = "Body Part Injured",
       y = "Number of Injuries for this Body Part",
       title = paste("Which Body Part was Most Injured in Ages", case_when(input$age_EL <= 10 ~ "0-10", 
                               (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                               (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                               (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                               (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                               (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                               (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                               (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75")),
       fill = "Body Part Injured") +
  interactive_theme

body_bar
}

})
```

### Data points from your specific age

```{r}
renderPlot({
  
body_polygon <- bdgramr(data = data, model = "basic_male") 

body_injury <- shark %>%
  mutate(body_1 = str_split_fixed(Injury.location, ", ", 4)[ ,1],
         body_2 = str_split_fixed(Injury.location, ", ", 4)[ ,2],
         body_3 = str_split_fixed(Injury.location, ", ", 4)[ ,3],
         body_4 = str_split_fixed(Injury.location, ", ", 4)[ ,4]) %>% 
    filter(Victim.age == as.numeric(input$age_EL))

if (input$gender_EL == "all") { #if the shark species is all
  
} else {
  body_injury <- body_injury %>% 
    filter(Victim.gender == input$gender_EL)
}

body_1 <- body_injury %>% 
  group_by(body_1) %>% 
  rename(body = body_1) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_2 <- body_injury %>% 
  group_by(body_2) %>% 
  rename(body = body_2) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_3 <- body_injury %>% 
  group_by(body_3) %>% 
  rename(body = body_3) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_4 <- body_injury %>% 
  group_by(body_4) %>% 
  rename(body = body_4) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_12 <- rbind(body_1, body_2)
body_123 <- rbind(body_12, body_3)
body_1234 <- rbind(body_123, body_4)

body_total <- body_1234 %>% 
  group_by(body) %>%
  summarize(Number_injury = sum(sum)) %>% 
  mutate(body = ifelse(body == "left leg", "leg", body),
         body = ifelse(body == "left foot", "foot", body),
         overall = case_when(body == "ankle" ~ "foot",
                            body == "calf" ~ "leg",
                            body == "hand" ~ "arm",
                            body == "leg" ~ "leg",
                            body == "other: half of body recovered" ~ "",
                            body == "other: uninjured" ~ "",
                            body == "right" ~ "",
                            body == "shoulder " ~ "shoulder",
                            body == "torso" ~ "torso",
                            body == "arm" ~ "arm",
                            body == "foot" ~ "foot",
                            body == "head" ~ "head",
                            body == "other: body not recovered" ~ "torso",
                            body == "other: multiple bites" ~ "torso",
                            body == "pelvic region" ~ "pelvic region",
                            body == "shoulder" ~ "shoulder",
                            body == "thigh" ~ "leg")) %>% 
  filter(overall != "") 

body_polygon <- bdgramr(data = data, model = "basic_male") %>% 
  mutate(body = case_when(Muscle == "Head" ~ "head",
                          Muscle == "Pectoral" ~ "shoulder",
                          Muscle == "Intercostal" ~ "torso",
                          Muscle == "Abdominal" ~ "pelvic region",
                          Muscle == "Quadriceps" ~ "thigh",
                          Muscle == "Foot" ~ "foot",
                          Muscle == "Shoulder" ~ "shoulder",
                          Muscle == "Lower_back" ~ "torso",
                          Muscle == "Hamstring" ~ "thigh",
                          Muscle == "Clavicle" ~ "neck",
                          Muscle == "Arm" ~ "arm",
                          Muscle == "Forearm" ~ "arm",
                          Muscle == "Hand" ~ "hand",
                          Muscle == "Shin" ~ "calf",
                          Muscle == "Neck" ~ "neck",
                          Muscle == "Upper_back" ~ "torso",
                          Muscle == "Gluteus_Maximus" ~ "pelvic region",
                          Muscle == "Calf" ~ "calf"))

body_combine <- full_join(body_polygon, body_total, by = "body", relationship = "many-to-many")


ggplot(data = body_combine, aes(x,y, group = Id)) +
  geom_bdgramr(color = "#67000d", aes(fill = Number_injury)) +
  scale_fill_gradient(low = "#fff5f0", high = "#a50f15" ) +
  labs(title = paste("Which body part was most injured at age", input$age_EL, "?"),
       caption = "BdygramR: original_male diagram",
       fill = "Number of Injuries") +
  interactive_theme

})
```

What Activities Put You at the Highest Risk of Shark Attacks?
=====================================  

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
radioButtons("gender_button", label = "Select Gender", choices = c("All" = "all", "Female" = "female", "Male" = "male"))
sliderInput("age_slider", "Select Age Range", min = 0, max = 90, value = c(0, 90))
```

Column
-----------------------------------------------------------------------

### What Activities Put You at the Highest Risk of Shark Attacks?

```{r}
sharks_act <- read.csv("Australian Shark-Incident Database Public Version.csv")

sharks_act <- sharks_act %>%
  select(Victim.activity, Victim.gender, Victim.age)

sharks_act <- sharks_act[-which(sharks_act$Victim.activity == ""), ]
sharks_act[sharks_act == "boarding"] <- "Boarding"
sharks_act[sharks_act == "diving"] <- "Diving"
sharks_act[sharks_act == "diving, collecting"] <- "Diving"
sharks_act[sharks_act == "fishing"] <- "Fishing"
sharks_act[sharks_act == "motorised boating"] <- "Boating"
sharks_act[sharks_act == "other: hull scraping"] <- "Other"
sharks_act[sharks_act == "other: jetskiing; swimming"] <- "Other"
sharks_act[sharks_act == "other: standing in water"] <- "Standing in Water"
sharks_act[sharks_act == "other:floating"] <- "Swimming"
sharks_act[sharks_act == "paddleboarding"] <- "Boarding"
sharks_act[sharks_act == "snorkeling"] <- "Snorkeling"
sharks_act[sharks_act == "snorkelling"] <- "Snorkeling"
sharks_act[sharks_act == "spearfishing"] <- "Spearfishing"
sharks_act[sharks_act == "surfing"] <- "Boarding"
sharks_act[sharks_act == "swimming"] <- "Swimming"
sharks_act[sharks_act == "unmotorised boating"] <- "Boating"
sharks_act[sharks_act == "wing foiler"] <- "Other"

renderPlotly({
  if(input$gender_button == "all")
    {
    sharks_act_filtered_v1 <- sharks_act
    }
  if(input$gender_button == "female")
    {
    sharks_act_filtered_v1 <- sharks_act %>%
      filter(Victim.gender == "female")
    }
  if(input$gender_button == "male")
    {
    sharks_act_filtered_v1 <- sharks_act %>%
      filter(Victim.gender == "male")
    }
  
  sharks_act_filtered <- sharks_act_filtered_v1 %>% filter(Victim.age %in% (as.numeric(input$age_slider[1]):as.numeric(input$age_slider[2])))
  
  p_act <- ggplot(sharks_act_filtered, aes(x = fct_infreq(Victim.activity), fill = Victim.activity)) + geom_bar() + scale_fill_manual(values = c("Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6")) + labs(x = "Activity", y = "Number of Shark Attacks") + interactive_theme
  
  ggplotly(p_act)})
```

When Should You Swim?
=====================================  

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
radioButtons("time_button", label = "By Month or Time of Day?", choices = c("Month" = "month", "Time of Day" = "hour"))
```

Column
-----------------------------------------------------------------------

### When Should You Swim?

```{r}
sharks_time <- read.csv("Australian Shark-Incident Database Public Version.csv")

sharks_time <- sharks_time %>%
  select(Incident.month, Time.of.incident, Victim.gender, Victim.age)

sharks_time$Time.of.incident <- floor(as.numeric(sharks_time$Time.of.incident) / 100) * 100

sharks_time_month <- sharks_time %>%
  group_by(Incident.month) %>%
  summarise(count_month = n())

sharks_time_month <- sharks_time_month %>%
  drop_na(Incident.month)

sharks_time_month$Incident.month <- factor(sharks_time_month$Incident.month, levels = 1:12, labels = month.name)

sharks_time_hour <- sharks_time %>%
  group_by(Time.of.incident) %>%
  summarise(count_hour = n())

for (n in seq(0, 2300, by = 100))
{
  if (!n %in% sharks_time_hour$Time.of.incident)
  {
    sharks_time_hour <- rbind(sharks_time_hour, c(n, 0))
  }
}

sharks_time_hour <- sharks_time_hour %>%
  drop_na(Time.of.incident)

renderPlotly({
  if(input$time_button == "month")
    {
    p_time <- ggplot(sharks_time_month, aes(x = Incident.month, y = count_month)) + geom_point(color = "#0a144a") + geom_line(group = 1, color = "#0a144a") + labs(x = "Month", y = "Number of Shark Attacks") + interactive_theme
    }
  else
    {
    p_time <- ggplot(sharks_time_hour, aes(x = Time.of.incident, y = count_hour)) + geom_point(color = "#0a144a") + geom_line(color = "#0a144a") + scale_x_continuous(breaks = seq(0, 2300, by = 200), labels = function(x) {
      hour <- floor(x / 100)
      minute <- x %% 100
      time <- sprintf("%02d:%02d", hour, minute)
      format(strptime(time, "%H:%M"), "%I:%M %p")
      }
      ) + labs(x = "Time of Day", y = "Number of Shark Attacks") + interactive_theme
    }
  ggplotly(p_time)
  })
```