---
title: "Shark Attacks in Australia"
date: "12/11/2024"
authors: "Erica Loomis (esloom25), Timothy Pratt (tjprat25), and Rebecca Wanger (rswang25)"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: 
      version: 4
      bg: "#101010"
      fg: "#8db8fc" 
      primary: "#8db8fc"
      base_font: 
        DIN Alternate
runtime: shiny
---

```{r}
## Connecting to the Shiny App Account
# rsconnect::setAccountInfo(name='ter1',
# 			  token='0163EF6118B243DE8D0042631C039930',
# 			  secret='<Secret>')

## Actually running and deploying the app

library(shiny)
library(ggplot2)
#runApp()

#deploying the app

library(rsconnect)
#deployApp()

#link: https://ter1.shinyapps.io/sc326-tre-interactive-graphics-project/
```

```{r global, include=FALSE, warning=FALSE, message=FALSE}
# loading in data in 'global' chunk so it can be shared by all users of the dashboard
shark_attacks <- read.csv("Australian Shark-Incident Database Public Version.csv")

# load packages in 'global' chunk so they can be shared by all users of the dashboard
library(httr2)
library(purrr)
library(tibble)
library(lubridate)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyr)
library(ggseas) #to use stat_rollapplyr
library(plotly)
library(forcats)

interactive_theme <- theme_gray() +# White background, black and white theme
  theme(plot.title = element_text(family = "Georgia", face = "bold", color = "#0a144a"),
        axis.title = element_text(family = "DIN Alternate", color = "#0a144a"),
        axis.text = element_text(color = "#0a144a"),
        legend.title = element_text(family = "DIN Alternate", face = "bold",  color = "#0a144a"),
        legend.text = element_text(family = "DIN Alternate", color = "#0a144a"),
        legend.box.margin = margin(6, 6, 6, 6),
        legend.background = element_rect(fill = "#d1d4e3"),
        panel.background = element_rect(fill = "#d1d4e3"),
        plot.background = element_rect(fill = "#d1d4e3"))
```

Introduction 
=====================================
```{r}
fluidPage(
  titlePanel("Australian Shark Attacks & Safety: An Interactive Analysis"),
  
  fluidRow(
    column(12,
      h3("Welcome to Australia!"),
      p("Shark attacks, while rare, are a significant cause for concern for beachgoers, swimmers, and surfers along Australia's coast. In this application, we analyze various factors contributing to shark attacks in Australia, including locations, timing, and shark species. Our goal is to help visitors and locals alike understand the patterns behind shark attacks and make informed decisions about water safety."),
      p("Through this interactive dashboard, you'll be able to explore historical data on shark attacks, view trends, and identify activities where safety measures may be most needed. Whether you're a tourist or a local, this tool aims to promote awareness and enhance safety practices when enjoying Australia's beautiful coastlines."),
      p("We hope this resource helps you have a safer and more informed experience in Australia!"),
      p("Click the buttons along the top bar to start exploring different graphs.")
    )
  ),
  
  fluidRow(
    column(6,
      h4("Data Source"),
      p("The data used in this application is sourced from the publicly available Australian Shark-Incident Database (ASID) published by the Taronga Conservation Society Australia."),
      p("The full dataset can be accessed here: https://taronga.org.au/conservation-and-science/australian-shark-incident-database#thedatabase")
    ),
    column(6,
      h4("Authors"),
      p("This application was developed by:"),
      tags$ul(
        tags$li("Erica Loomis"),
        tags$li("Timothy Pratt"),
        tags$li("Rebecca Wanger"),
      )
    )
  )
)


```

How fatal is this shark? {data-navmenu="Menu"}
=====================================  

```{r global_species, include=FALSE, warning=FALSE, message=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(tidyverse)
library(flexdashboard)
library(plotly)
library(stringr)
library(knitr)
library(dplyr)


shark <- read.csv("Australian Shark-Incident Database Public Version.csv")
blurb <- read.csv("shark blurb.csv")


```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("shark_species_EL", label = "Attacks based on shark species", 
            choices =  c("All" = "all",
                         "White shark" = "white shark",
                         "Tiger shark" = "tiger shark",
                         "Wobbegong" = "wobbegong",
                         "Bull shark" = "bull shark",
                         "Whaler shark" = "whaler shark",
                         "Bronze whaler shark" = "bronze whaler shark",
                         "Grey nurse shark" = "grey nurse shark",
                         "Grey reef shark" = "grey reef shark",
                         "Whitetip reef shark" = "whitetip reef shark",
                         "Lemon shark" = "lemon shark",
                         "Hammerhead shark" = "hammerhead shark ",
                         "Dusky shark" = "dusky shark",
                         "Blacktip reef shark" = "blacktip reef shark",
                         "Galapagos shark" = "galapagos shark",
                         "Shortfin mako shark" = "shortfin mako shark",
                         "Broadnose sevengill shark" = "broadnose sevengill shark",
                         "Silvertip shark" = "silvertip shark",
                         "Sevengill shark" = "sevengill shark",
                         "School shark" = "school shark",
                         "Port jackson shark" = "port jackson shark",
                         "Dogfish" = "dogfish",
                         "Blind shark" = "blind shark"))

selectInput("provoke_EL", label = "Attacks if provoked or unprovoked", 
            choices = c("Provoked and Unprovoked" = "both",
                        "Provoked" = "provoked",
                        "Unprovoked" = "unprovoked"))
```

```{r}
 renderTable({
    blurb %>%
    filter(name == input$shark_species_EL) %>% 
    select(Takeaways)

})

renderImage({
    if(input$shark_species_EL == "all"){
      list(
          src = file.path(paste("shark.jpg")),
          contentType = "image/jpeg",
          width = 230,
          alt = "Illustration of a shark"
        )
    } else {
        list(
          src = file.path(paste(input$shark_species_EL, ".jpg")),
          contentType = "image/jpeg",
          width = 230,
          alt = paste("Picture of a", input$shark_species_EL)
        )
    }
  }, deleteFile = FALSE)

```

Row
-----------------------------------------------------------------------
### Percent Fatality

```{r}
fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup()

renderGauge({
  
if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      mutate(total_attacks = sum(sum_species),
             percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))

  } else { # and they either provoke or unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      mutate(total_attacks = sum(sum_species),
           percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } 
} else { # if is is a specific shark species
  if(input$provoke_EL == "both") { #and they provoke and unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species),
             percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
      group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } else { # and they either provoke or unprovoke
    fatal_percent_gauge <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species),
           percent_fatal = (sum_species / total_attacks) * 100) %>% 
      filter(fatal == "fatal") %>% 
        group_by(total_attacks) %>% 
      summarize(percent_fatal = sum(percent_fatal))
  } 
}
  rate <- fatal_percent_gauge$percent_fatal # assign value 
  gauge(rate, min = 0, max = 100, symbol = '%', gaugeSectors( # render the gauge
    success = c(0, 39), warning = c(40, 79), danger = c(80, 100)
  ))
})
```

### Number of Attacks in this Situation
```{r}
fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup()

renderValueBox({
  
  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
}
    if(is.nan(fatal_percent_value$total_attacks)){
    attacks <- 0
  } else {
    attacks <- fatal_percent_value$total_attacks #assign value
  }
  valueBox(attacks, # render the valuebox
           icon = "ion-alert-circled",
           color = "#0a144a")
})
```

### Number of Deaths
```{r}

fatal_percent <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>% 
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>% 
  summarize(sum_species = n()) %>% 
  ungroup() %>% 
  filter( Shark.common.name != "") 

renderValueBox({
  
  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Shark.common.name == input$shark_species_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } else { # and they either provoke or unprovoke
    fatal_percent_value <- fatal_percent %>% 
      filter(Provoked.unprovoked == input$provoke_EL, Shark.common.name ==
               input$shark_species_EL) %>% 
      filter(fatal == "fatal") %>% 
      mutate(total_attacks = sum(sum_species)) %>% 
      summarize(total_attacks = mean(total_attacks))
  } 
}
  if(is.nan(fatal_percent_value$total_attacks)){
    deaths <- 0
  } else {
    deaths <- fatal_percent_value$total_attacks #assign value
  }
  valueBox(deaths, # render the valuebox
           icon = "ion-alert-circled",
           color = "#67000d")
})
```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### All species

```{r}
  shark_bubble <- shark %>%
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
  group_by(Shark.common.name, Provoked.unprovoked, fatal) %>%
  summarize(attack_species_provoke_fatal = n()) %>%
  mutate(species_provoke = paste(Shark.common.name, Provoked.unprovoked, sep=" ")) %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  mutate(attack_species_provoke = sum(attack_species_provoke_fatal)) %>%
  group_by(Shark.common.name, fatal) %>%
  mutate(attack_species_fatal = sum(attack_species_provoke_fatal)) %>%
  group_by(Shark.common.name) %>%
  mutate(attack_species = sum(attack_species_provoke_fatal), # all attacks by species
         all_percent = ifelse(fatal == "fatal", attack_species_fatal / attack_species, 0), # calculating the percent fatal for all
         provoke_percent = ifelse(fatal == "fatal", attack_species_provoke_fatal / attack_species_provoke, 0)) %>%
  filter( Shark.common.name != "") %>%
  mutate(Shark.common.name = str_to_title(Shark.common.name))


renderPlotly({ #going into the if statements

  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    
    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent) * 100) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point(size = 3) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } else { # and they either provoke or unprovoke
    
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent) * 100) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point(size = 3) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = paste("Which Shark Species Attack is Most Common and Fatal?-", input$provoke_EL)) +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph

  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke

    shark_final <- shark_bubble %>% 
  group_by(Shark.common.name) %>% #only grouping by shark species
  summarize(attack_species = max(attack_species), percent = max(all_percent) * 100) %>% 
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
  
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } else { # and they either provoke or unprovoke
    shark_final <- shark_bubble %>%
  group_by(Shark.common.name, Provoked.unprovoked) %>%
  summarize(attack_species = max(attack_species_provoke), percent = max(provoke_percent) * 100) %>% 
        filter(Provoked.unprovoked == input$provoke_EL) %>% # only taking the provoked or unprovoked options
  mutate(text = paste("Shark Name: ", Shark.common.name, "\nNumber of Attacks: ", attack_species, "\nFatal attacks: ", round(percent, digits = 1), "%"))
    
shark_final_highlight <- shark_final %>% 
  filter(Shark.common.name == str_to_title(input$shark_species_EL))
    
bubble <- ggplot(shark_final, aes(x= attack_species, y= reorder(Shark.common.name, attack_species), color = percent, text = text)) +
  geom_point() +
  geom_point(data = shark_final_highlight, size = 5) +
  scale_colour_gradient(low = "#fcbba1", high = "#a50f15") +
  labs(x = "Number of Attacks per Species",
       y = "Shark Common Names",
       color = "Percent fatality",
       title = "Which Shark Species Attack is Most Common and Fatal?") +
  interactive_theme

ggplotly(bubble, tooltip="text") #plotly graph
    
  } 
}


})
```


### Shark Length
```{r}
renderPlot({ #going into the if statements

  if (input$shark_species_EL == "all") { #if the shark species is all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke
    
    fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal"))
    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
  scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
    
  } else { # and they either provoke or unprovoke
    
  fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
    filter(Provoked.unprovoked == input$provoke_EL) # only taking the provoked or unprovoked options

  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  } 
} else { # if the shark species is not all
  if(input$provoke_EL == "both") { # and they both provoke and unprovoke

  fatal_shark <- shark %>% 
    mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
      filter(Shark.common.name == input$shark_species_EL)


    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  
  } else { # and they either provoke or unprovoke
    
fatal_shark <- shark %>% 
  mutate(fatal = ifelse(Victim.injury == "fatal", "fatal", "not fatal")) %>%
    filter(Provoked.unprovoked == input$provoke_EL) %>%  # only taking the provoked or unprovoked options
    filter(Shark.common.name == input$shark_species_EL)


    
  ggplot(fatal_shark, aes(fatal, Shark.length.m, fill = fatal)) +
  geom_boxplot( ) +
      scale_fill_manual(values = c("#8c3a40", "#343d6b")) +
  scale_color_manual(values = c("#8c3a40","#343d6b")) +
  geom_rug( aes(color = fatal)) +
  theme(legend.position = "none") +
  labs(x = "Fatal or Non-Fatal Attack",
       y = "Shark Length (m)",
       title = "Are Longer Sharks More Fatal?") +
    interactive_theme
  } 
}


})
```

Which body part are you most likely to injure? {data-navmenu="Menu"}
=====================================  

```{r global_body, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
library(stringr)
library(devtools)
library(bdgramR)

shark <- read.csv("Australian Shark-Incident Database Public Version.csv")

```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("gender_EL", label = "Select a gender", 
            choices =  c("All" = "all",
                         "Male" = "male",
                         "Female" = "female"))

selectInput("age_EL", label = "Select an age range", 
            choices =  c("0-10" = "0-10",
                         "11-20" = "11-20",
                         "21-30" = "21-30",
                         "31-40" = "31-40",
                         "41-50" = "41-50",
                         "51-60" = "51-60",
                         "61-70" = "61-70",
                         "71-75" = "71-75"))

radioButtons("bar_EL", "Bar Chart type",
               c("Bar chart of all age ranges" = "all",
                 "Bar chart of your age range" = "specific"))


```


Row {data-height=130}
-----------------------------------------------------------------------

### Helpful Advice for your trip
During your Australia trip, we would love for you to be as safe as possible. On this page we visualized some data representing where people have been injured in the past due to shark attacks so you can watch out for said body part depending on gender and age range. In general, people between the age 10 and 40 were injured most on their legs, and people between the ages 41 to 60 were injured most on their arms. People seem to also get injured the most on their legs regardless of what gender they were. Keeps those legs close to you!

Row 
-----------------------------------------------------------------------
### Body part injuries by age chunk

```{r}
renderPlot({
  
body_injury <- shark %>%
  mutate(body_1 = str_split_fixed(Injury.location, ", ", 4)[ ,1],
         body_2 = str_split_fixed(Injury.location, ", ", 4)[ ,2],
         body_3 = str_split_fixed(Injury.location, ", ", 4)[ ,3],
         body_4 = str_split_fixed(Injury.location, ", ", 4)[ ,4],
         age_range = case_when(Victim.age <= 10 ~ "0-10", 
                              (10 < Victim.age) & (Victim.age <= 20)  ~ "11-20",
                              (20 < Victim.age) & (Victim.age <= 30)  ~ "21-30",
                              (30 < Victim.age) & (Victim.age <= 40)  ~ "31-40",
                              (40 < Victim.age) & (Victim.age <= 50)  ~ "41-50",
                              (50 < Victim.age) & (Victim.age <= 60)  ~ "51-60",
                              (60 < Victim.age) & (Victim.age <= 70)  ~ "61-70",
                              (70 < Victim.age) & (Victim.age <= 75)  ~ "71-75",
                              Victim.age == " "                ~ NA)) 
if (input$gender_EL == "all") { #if it's all genders
  
} else { #if it's a specific gender
  body_injury <- body_injury %>% 
    filter(Victim.gender == input$gender_EL)
}

body_1 <- body_injury %>% 
  group_by(body_1, age_range) %>% 
  rename(body = body_1) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_2 <- body_injury %>% 
  group_by(body_2, age_range) %>% 
  rename(body = body_2) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_3 <- body_injury %>% 
  group_by(body_3, age_range) %>% 
  rename(body = body_3) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_4 <- body_injury %>% 
  group_by(body_4, age_range) %>% 
  rename(body = body_4) %>% 
  filter(body != "") %>% 
  summarize(sum = n())

body_12 <- rbind(body_1, body_2)
body_123 <- rbind(body_12, body_3)
body_1234 <- rbind(body_123, body_4)

body_total <- body_1234 %>% 
  group_by(body, age_range) %>%
  mutate(body = ifelse(body == "left leg", "leg", body),
         body = ifelse(body == "left foot", "foot", body)) %>% 
  mutate(overall = case_when(body == "ankle" ~ "foot",
                             body == "calf" ~ "leg",
                             body == "hand" ~ "arm",
                             body == "leg" ~ "leg",
                             body == "other: half of body recovered" ~ "",
                             body == "other: uninjured" ~ "",
                             body == "right" ~ "",
                             body == "shoulder " ~ "shoulder",
                             body == "torso" ~ "torso",
                             body == "arm" ~ "arm",
                             body == "foot" ~ "foot",
                             body == "head" ~ "head",
                             body == "other: body not recovered" ~ "",
                             body == "other: multiple bites" ~ "",
                             body == "pelvic region" ~ "pelvic region",
                             body == "shoulder" ~ "shoulder",
                             body == "thigh" ~ "leg"),
         body = ifelse(grepl("other", body, fixed = TRUE), "", body)) %>% 
  filter(overall != "" ) %>% 
  group_by(overall, age_range) %>% 
  summarize(Number_injury = sum(sum)) %>% 
  na.omit()

final_body_bar <- body_total %>% 
    group_by(age_range) %>% 
  filter(Number_injury == max(Number_injury))

final_body_bar$age_range <- factor(final_body_bar$age_range, levels = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71-75"))

body_bar <- ggplot(final_body_bar, aes(x = age_range, y = Number_injury, fill = overall)) +
  geom_col() +
  labs(x = "Age Range",
       y = "Number of Injuries for this Body Part",
       title = "Which Body Part was Most Injured by Age Ranges?",
       fill = "Body Part Injured") +
  scale_fill_manual(values = c("#44AA99", "#CC6677", "#493C86", "#DDCC77")) 
if (input$bar_EL == "all") { #if it's all barchart
    
 specific_age <- final_body_bar %>% 
  filter(age_range == case_when(input$age_EL <= 10 ~ "0-10", 
                              (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                              (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                              (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                              (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                              (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                              (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                              (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75"))

body_bar + geom_col(data = specific_age, color = "#a50f15", size = 1.5) + 
  guides(color = "none") +
  interactive_theme
  
} else { #if it's a specific age barchart

body_total_age <- body_total %>% 
  group_by(overall) %>% 
  filter(age_range == case_when(input$age_EL <= 10 ~ "0-10", 
                               (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                               (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                               (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                               (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                               (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                               (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                               (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75"))

body_bar <- ggplot(body_total_age, aes(x = reorder(overall, Number_injury), y = Number_injury)) +
  geom_col(fill = "#a50f15") +
  labs(x = "Body Part Injured",
       y = "Number of Injuries for this Body Part",
       title = paste("Which Body Part was Most Injured in Ages", case_when(input$age_EL <= 10 ~ "0-10", 
                               (10 < input$age_EL) & (input$age_EL <= 20)  ~ "11-20",
                               (20 < input$age_EL) & (input$age_EL <= 30)  ~ "21-30",
                               (30 < input$age_EL) & (input$age_EL <= 40)  ~ "31-40",
                               (40 < input$age_EL) & (input$age_EL <= 50)  ~ "41-50",
                               (50 < input$age_EL) & (input$age_EL <= 60)  ~ "51-60",
                               (60 < input$age_EL) & (input$age_EL <= 70)  ~ "61-70",
                               (70 < input$age_EL) & (input$age_EL <= 75)  ~ "71-75")),
       fill = "Body Part Injured") +
  interactive_theme 

body_bar
}

})
```

### Which body part is most likely to get injured?

```{r}
renderPlot({
  
 body_polygon <- bdgramr(data = data, model = "basic_male") 
  
  body_injury <- shark %>%
    mutate(body_1 = str_split_fixed(Injury.location, ", ", 4)[ ,1],
           body_2 = str_split_fixed(Injury.location, ", ", 4)[ ,2],
           body_3 = str_split_fixed(Injury.location, ", ", 4)[ ,3],
           body_4 = str_split_fixed(Injury.location, ", ", 4)[ ,4],
           age_range = case_when(Victim.age <= 10 ~ "0-10", 
                                 (10 < Victim.age) & (Victim.age <= 20)  ~ "11-20",
                                 (20 < Victim.age) & (Victim.age <= 30)  ~ "21-30",
                                 (30 < Victim.age) & (Victim.age <= 40)  ~ "31-40",
                                 (40 < Victim.age) & (Victim.age <= 50)  ~ "41-50",
                                 (50 < Victim.age) & (Victim.age <= 60)  ~ "51-60",
                                 (60 < Victim.age) & (Victim.age <= 70)  ~ "61-70",
                                 (70 < Victim.age) & (Victim.age <= 75)  ~ "71-75",
                                 Victim.age == " "                ~ NA)) 
  
  if (input$gender_EL == "all") { #if it's all genders

  } else { #if it's a specific gender
    body_injury <- body_injury %>%
      filter(Victim.gender == input$gender_EL)
  }
  
  body_1 <- body_injury %>% 
    group_by(body_1, age_range) %>% 
    rename(body = body_1) %>% 
    filter(body != "") %>% 
    summarize(sum = n())
  
  body_2 <- body_injury %>% 
    group_by(body_2, age_range) %>% 
    rename(body = body_2) %>% 
    filter(body != "") %>% 
    summarize(sum = n())
  
  body_3 <- body_injury %>% 
    group_by(body_3, age_range) %>% 
    rename(body = body_3) %>% 
    filter(body != "") %>% 
    summarize(sum = n())
  
  body_4 <- body_injury %>% 
    group_by(body_4, age_range) %>% 
    rename(body = body_4) %>% 
    filter(body != "") %>% 
    summarize(sum = n())
  
  body_12 <- rbind(body_1, body_2)
  body_123 <- rbind(body_12, body_3)
  body_1234 <- rbind(body_123, body_4)
  
  body_total <- body_1234 %>% 
    group_by(body, age_range) %>%
    mutate(body = ifelse(body == "left leg", "leg", body),
           body = ifelse(body == "left foot", "foot", body)) %>% 
    mutate(overall = case_when(body == "ankle" ~ "foot",
                               body == "calf" ~ "leg",
                               body == "hand" ~ "arm",
                               body == "leg" ~ "leg",
                               body == "other: half of body recovered" ~ "",
                               body == "other: uninjured" ~ "",
                               body == "right" ~ "",
                               body == "shoulder " ~ "shoulder",
                               body == "torso" ~ "torso",
                               body == "arm" ~ "arm",
                               body == "foot" ~ "foot",
                               body == "head" ~ "head",
                               body == "other: body not recovered" ~ "",
                               body == "other: multiple bites" ~ "",
                               body == "pelvic region" ~ "pelvic region",
                               body == "shoulder" ~ "shoulder",
                               body == "thigh" ~ "leg"),
           body = ifelse(grepl("other", body, fixed = TRUE), "", body)) %>% 
    filter(overall != "" ) %>% 
    group_by(overall, age_range) %>% 
    summarize(Number_injury = sum(sum)) %>% 
    na.omit() %>% 
    filter(age_range == input$age_EL)
  
  
  body_polygon <- bdgramr(data = data, model = "basic_male") %>% 
    mutate(body = case_when(Muscle == "Head" ~ "head",
                            Muscle == "Pectoral" ~ "shoulder",
                            Muscle == "Intercostal" ~ "torso",
                            Muscle == "Abdominal" ~ "pelvic region",
                            Muscle == "Quadriceps" ~ "leg",
                            Muscle == "Foot" ~ "foot",
                            Muscle == "Shoulder" ~ "shoulder",
                            Muscle == "Lower_back" ~ "torso",
                            Muscle == "Hamstring" ~ "leg",
                            Muscle == "Clavicle" ~ "neck",
                            Muscle == "Arm" ~ "arm",
                            Muscle == "Forearm" ~ "arm",
                            Muscle == "Hand" ~ "hand",
                            Muscle == "Shin" ~ "leg",
                            Muscle == "Neck" ~ "neck",
                            Muscle == "Upper_Back" ~ "torso",
                            Muscle == "Gluteus_Maximus" ~ "pelvic region",
                            Muscle == "Calf" ~ "leg"))
  
  body_combine <- full_join(body_polygon, body_total, by = c("body" = "overall"), relationship = "many-to-one")


ggplot(data = body_combine, aes(x,y, group = Id)) +
  geom_bdgramr(color = "#67000d", aes(fill = Number_injury)) +
  scale_fill_gradient(low = "#fff5f0", high = "#a50f15" ) +
  labs(title = paste("Which body part was most injured at age", input$age_EL, "?"),
       caption = "BdygramR: original_male diagram",
       fill = "Number of Injuries") +
  interactive_theme

})
```

What Activities Put You at the Highest Risk of Shark Attacks? {data-navmenu="Menu"}
=====================================  

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
radioButtons("gender_button", label = "Select Gender", choices = c("All" = "all", "Female" = "female", "Male" = "male"))
sliderInput("age_slider", "Select Age Range", min = 0, max = 90, value = c(0, 90))
```

Column
-----------------------------------------------------------------------

### What Activities Put You at the Highest Risk of Shark Attacks?

```{r}
sharks_act <- read.csv("Australian Shark-Incident Database Public Version.csv")

sharks_act <- sharks_act %>%
  select(Victim.activity, Victim.gender, Victim.age)

sharks_act <- sharks_act[-which(sharks_act$Victim.activity == ""), ]
sharks_act[sharks_act == "boarding"] <- "Boarding"
sharks_act[sharks_act == "diving"] <- "Diving"
sharks_act[sharks_act == "diving, collecting"] <- "Diving"
sharks_act[sharks_act == "fishing"] <- "Fishing"
sharks_act[sharks_act == "motorised boating"] <- "Boating"
sharks_act[sharks_act == "other: hull scraping"] <- "Other"
sharks_act[sharks_act == "other: jetskiing; swimming"] <- "Other"
sharks_act[sharks_act == "other: standing in water"] <- "Standing in Water"
sharks_act[sharks_act == "other:floating"] <- "Swimming"
sharks_act[sharks_act == "paddleboarding"] <- "Boarding"
sharks_act[sharks_act == "snorkeling"] <- "Snorkeling"
sharks_act[sharks_act == "snorkelling"] <- "Snorkeling"
sharks_act[sharks_act == "spearfishing"] <- "Spearfishing"
sharks_act[sharks_act == "surfing"] <- "Boarding"
sharks_act[sharks_act == "swimming"] <- "Swimming"
sharks_act[sharks_act == "unmotorised boating"] <- "Boating"
sharks_act[sharks_act == "wing foiler"] <- "Other"

renderPlotly({
  if(input$gender_button == "all")
    {
    sharks_act_filtered_v1 <- sharks_act
    }
  if(input$gender_button == "female")
    {
    sharks_act_filtered_v1 <- sharks_act %>%
      filter(Victim.gender == "female")
    }
  if(input$gender_button == "male")
    {
    sharks_act_filtered_v1 <- sharks_act %>%
      filter(Victim.gender == "male")
    }
  
  sharks_act_filtered <- sharks_act_filtered_v1 %>% filter(Victim.age %in% (as.numeric(input$age_slider[1]):as.numeric(input$age_slider[2])))
  
  p_act <- ggplot(sharks_act_filtered, aes(x = fct_infreq(Victim.activity), fill = Victim.activity)) + geom_bar() + scale_fill_manual(values = c("Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6")) + labs(x = "Activity", y = "Number of Shark Attacks") + interactive_theme
  
  ggplotly(p_act)})
```

When Should You Swim? {data-navmenu="Menu"}
=====================================  

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
radioButtons("time_button", label = "By Month or Time of Day?", choices = c("Month" = "month", "Time of Day" = "hour"))
```

Column
-----------------------------------------------------------------------

### When Should You Swim?

```{r}
sharks_time <- read.csv("Australian Shark-Incident Database Public Version.csv")

sharks_time <- sharks_time %>%
  select(Incident.month, Time.of.incident, Victim.gender, Victim.age)

sharks_time$Time.of.incident <- floor(as.numeric(sharks_time$Time.of.incident) / 100) * 100

sharks_time_month <- sharks_time %>%
  group_by(Incident.month) %>%
  summarise(count_month = n())

sharks_time_month <- sharks_time_month %>%
  drop_na(Incident.month)

sharks_time_month$Incident.month <- factor(sharks_time_month$Incident.month, levels = 1:12, labels = month.name)

sharks_time_hour <- sharks_time %>%
  group_by(Time.of.incident) %>%
  summarise(count_hour = n())

for (n in seq(0, 2300, by = 100))
{
  if (!n %in% sharks_time_hour$Time.of.incident)
  {
    sharks_time_hour <- rbind(sharks_time_hour, c(n, 0))
  }
}

sharks_time_hour <- sharks_time_hour %>%
  drop_na(Time.of.incident)

renderPlotly({
  if(input$time_button == "month")
    {
    p_time <- ggplot(sharks_time_month, aes(x = Incident.month, y = count_month)) + geom_point(color = "#0a144a") + geom_line(group = 1, color = "#0a144a") + labs(x = "Month", y = "Number of Shark Attacks") + interactive_theme
    }
  else
    {
    p_time <- ggplot(sharks_time_hour, aes(x = Time.of.incident, y = count_hour)) + geom_point(color = "#0a144a") + geom_line(color = "#0a144a") + scale_x_continuous(breaks = seq(0, 2300, by = 200), labels = function(x) {
      hour <- floor(x / 100)
      minute <- x %% 100
      time <- sprintf("%02d:%02d", hour, minute)
      format(strptime(time, "%H:%M"), "%I:%M %p")
      }
      ) + labs(x = "Time of Day", y = "Number of Shark Attacks") + interactive_theme
    }
  ggplotly(p_time)
  })
```

How Have Shark Attacks Changed Over Time? {data-navmenu="Menu"}
=====================================  

```{r global_time, include=FALSE, warning=FALSE, message=FALSE}

# load packages in 'global' chunk so they can be shared by all users of the dashboard
library(httr2)
library(purrr)
library(tibble)
library(lubridate)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyr)
library(ggseas) #to use stat_rollapplyr
library(plotly)

#loading in packages related to the maps
library(stringr) #to use str functions
library(sf)
library(maps)

# Loading in the Data and Packages

shark_attacks <- read.csv("Australian Shark-Incident Database Public Version.csv")

# reading in the Australian state shapefiles

load("aus_shp.RData")
  #st_read("./STE_2021_AUST_SHP_GDA94/STE_2021_AUST_GDA94.shp")



```


```{r}

# cleaning the data

#adding in means, renaming and recategorizing, 
shark_attacks <- shark_attacks %>%
  filter(is.na(UIN) != T) %>% #removing the empty rows
  mutate(Victim.injury = case_when(Victim.injury == "fatal" ~ "Fatal",
                                   Victim.injury == "injured" ~ "Injured",
                                   Victim.injury == "Injured" ~ "Injured",
                                   Victim.injury == "uninjured" ~ "Uninjured")) %>%
  mutate(Provoked.unprovoked = case_when(Provoked.unprovoked == "provoked" ~ "Provoked",
                                   Provoked.unprovoked == "unprovoked" ~ "Unprovoked")) %>%
  mutate(Victim.gender = case_when(Victim.gender == "female" ~ "Female",
                                   Victim.gender == "male" ~ "Male")) %>%
  mutate(Injury.severity = case_when(Injury.severity == "abrasion" ~ "Abrasion", Injury.severity == "major lacerations" ~ "Major Lacerations", Injury.severity == "minor lacerations" ~ "Minor Lacerations", Injury.severity == "other: teeth marks" ~ "Other: Teeth Marks", Injury.severity == "punctures" ~ "Punctures", Injury.severity == "surface wound" ~ "Surface Wound")) %>%
  mutate(Site.category = case_when(Site.category == "coastal" ~ "Coastal",
                                   Site.category == "Coastal" ~ "Coastal",
                                   Site.category == "estuary/harbour" ~ "Estuary/Harbour",
                                   Site.category == "island open ocean" ~ "Island Open Ocean", 
                                   Site.category == "ocean/pelagic" ~ "Ocean/Pelagic",
                                   Site.category == "other: fish farm" ~ "Other: Fish Farm",
                                   Site.category == "river" ~ "River")) %>%
  filter(Incident.year != "1791") %>%
  mutate(Victim.activity = case_when(Victim.activity == "boarding" ~ "Boarding",
                                   Victim.activity == "diving" ~ "Diving",
                                   Victim.activity == "diving, collecting" ~ "Diving",
                                   Victim.activity == "fishing" ~ "Fishing",
                                   Victim.activity == "motorised boating" ~ "Motorised Boating",
                                   Victim.activity == "other: standing in water" ~ "Standing in Water",
                                   Victim.activity == "other: hull scraping" ~ "Other",
                                   Victim.activity == "other: jetskiing; swimming" ~ "Other",
                                   Victim.activity == "other:floating" ~ "Swimming",
                                   Victim.activity == "wing foiler" ~ "Other",
                                   Victim.activity == "snorkeling" ~ "Snorkeling", 
                                   Victim.activity == "surfing" ~ "Boarding",
                                   Victim.activity == "snorkelling" ~ "Snorkeling",
                                   Victim.activity == "spearfishing" ~ "Spearfishing",
                                   Victim.activity == "paddleboarding" ~ "Boarding",
                                   Victim.activity == "swimming" ~ "Swimming",
                                   Victim.activity == "unmotorised boating" ~ "Boating",
                                   Victim.activity == "" ~ NA)) %>%
  filter(Incident.year != "1791") %>%
  group_by(Incident.year) %>%
   #creating mean columns
  mutate(Victim.age.mean = mean(Victim.age)) %>%
  #creating date groupings
  mutate(Year = case_when((Incident.year > 1800 & Incident.year <= 1850) ~ "1800-1850", 
                          (Incident.year > 1850 & Incident.year <= 1900) ~ "1850-1900",
                          (Incident.year > 1900 & Incident.year <= 1950) ~ "1900-1950",
                          (Incident.year > 1950 & Incident.year <= 2000) ~ "1950-2000",
                          (Incident.year > 2000 & Incident.year <= 2024) ~ "2000-2024")) 

#counting the number of sharks every 50 years
shark_attacks <- shark_attacks %>%
  group_by(Year) %>%
  mutate(Number_Attacks = n())
```

```{r}
#creating a dataframe for timeline data
shark_attacks_timeline <- shark_attacks %>%
  mutate(Victim.age = as.ts(Victim.age)) %>% 
  mutate(Incident.month = as.character(Incident.month)) %>% 
  mutate(Incident.year = as.character(Incident.year)) %>%
  mutate(Incident.year = make_date(year = Incident.year))
  # unite("Incident.month_year", Incident.month, Incident.year, sep = "/", remove = F) %>% #remove = F, na.rm = T
  # mutate(Incident.month_year = paste0("01/", Incident.month_year),
  #        Incident.month_year = as.Date(Incident.month_year, format = "%d/%m/%Y"))
  # 
#start and stop times

start_time <- as.Date("1900-01-01")
end_time <- as.Date("2024-03-01")

start.end <- c(start_time,end_time)

shark_attacks_count <- shark_attacks_timeline %>%
  group_by(Incident.year) %>%
  summarise(attack_count = n()) %>%
  mutate(Incident.year = as.Date(Incident.year, format = "%Y"))

```

```{r}

# computing proportions individually

#individually

#Victim.injury
shark_attacks_proportions_Victim.injury <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.injury != "") %>%
  group_by(Year, Victim.injury) %>%  #  group by the year variable
  summarise(count_Victim.injury = n()) %>% 
  mutate(total_Victim.injury = sum(count_Victim.injury),  #  compute total number of observations
         proportion_Victim.injury = count_Victim.injury / total_Victim.injury,  #  compute proportions
         std_error_Victim.injury = sqrt(proportion_Victim.injury * (1-proportion_Victim.injury) / total_Victim.injury),  #  compute standard error of each proportion
         lower_Victim.injury = proportion_Victim.injury - 1.96 * std_error_Victim.injury,  #  compute CI lower bound
         upper_Victim.injury = proportion_Victim.injury + 1.96 * std_error_Victim.injury) 
  

#Site.category
shark_attacks_proportions_Site.category <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Site.category != "") %>%
  group_by(Year, Site.category) %>%  #  group by the year variable
  summarise(count_Site.category = n()) %>%  
  mutate(total_Site.category = sum(count_Site.category),  
         proportion_Site.category = count_Site.category / total_Site.category,  
         std_error_Site.category = sqrt(proportion_Site.category * (1-proportion_Site.category) / total_Site.category),  
         lower_Site.category = proportion_Site.category - 1.96 * std_error_Site.category,  
         upper_Site.category = proportion_Site.category + 1.96 * std_error_Site.category)
  

#Provoked.unprovoked
shark_attacks_proportions_Provoked.unprovoked <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Provoked.unprovoked != "") %>%
  group_by(Year, Provoked.unprovoked) %>%  #  group by the year variable
  summarise(count_Provoked.unprovoked = n()) %>%  
  mutate(total_Provoked.unprovoked = sum(count_Provoked.unprovoked),  
         proportion_Provoked.unprovoked = count_Provoked.unprovoked / total_Provoked.unprovoked, 
         std_error_Provoked.unprovoked = sqrt(proportion_Provoked.unprovoked * (1-proportion_Provoked.unprovoked) / total_Provoked.unprovoked),  
         lower_Provoked.unprovoked = proportion_Provoked.unprovoked - 1.96 * std_error_Provoked.unprovoked,  
         upper_Provoked.unprovoked = proportion_Provoked.unprovoked + 1.96 * std_error_Provoked.unprovoked)
  

#Victim.activity
shark_attacks_proportions_Victim.activity <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.activity != "") %>%
  group_by(Year, Victim.activity) %>%  #  group by the year variable
  summarise(count_Victim.activity = n()) %>%  
  mutate(total_Victim.activity = sum(count_Victim.activity),  
         proportion_Victim.activity = count_Victim.activity / total_Victim.activity,  
         std_error_Victim.activity = sqrt(proportion_Victim.activity * (1-proportion_Victim.activity) / total_Victim.activity),  
         lower_Victim.activity = proportion_Victim.activity - 1.96 * std_error_Victim.activity,  
         upper_Victim.activity = proportion_Victim.activity + 1.96 * std_error_Victim.activity)


#Victim.gender 
shark_attacks_proportions_Victim.gender <- shark_attacks %>% #  Start with the data.frame
  filter(Victim.gender != "") %>%
  group_by(Year, Victim.gender) %>%  #  group by the year variable
  summarise(count_Victim.gender = n()) %>%    
  mutate(total_Victim.gender = sum(count_Victim.gender),  
         proportion_Victim.gender = count_Victim.gender / total_Victim.gender,  
         std_error_Victim.gender = sqrt(proportion_Victim.gender * (1-proportion_Victim.gender) / total_Victim.gender), 
         lower_Victim.gender = proportion_Victim.gender - 1.96 * std_error_Victim.gender,  
         upper_Victim.gender = proportion_Victim.gender + 1.96 * std_error_Victim.gender)
  
  
#Injury.severity
shark_attacks_proportions_Injury.severity <- shark_attacks %>%  #Start with the cereal data.frame
  filter(Injury.severity != "") %>%
  group_by(Year, Injury.severity) %>%  #  group by the year variable
  summarise(count_Injury.severity = n()) %>%  
  mutate(total_Injury.severity = sum(count_Injury.severity),  
         proportion_Injury.severity = count_Injury.severity / total_Injury.severity,  
         std_error_Injury.severity = sqrt(proportion_Injury.severity * (1-proportion_Injury.severity) / total_Injury.severity), 
         lower_Injury.severity = proportion_Injury.severity - 1.96 * std_error_Injury.severity,  
         upper_Injury.severity = proportion_Injury.severity + 1.96 * std_error_Injury.severity) 





```

### Shark Attacks over Time



```{r}


dataframe_barchart <- reactive({switch(input$response_var,
                    "Victim.injury" = shark_attacks_proportions_Victim.injury,
                    "Site.category" = shark_attacks_proportions_Site.category,
                    "Provoked.unprovoked" = shark_attacks_proportions_Provoked.unprovoked,
                    "Victim.activity" = shark_attacks_proportions_Victim.activity,
                    "Injury.severity" = shark_attacks_proportions_Injury.severity,
                     "Victim.gender" = shark_attacks_proportions_Victim.gender)
})

dataframe_time <- reactive({switch(input$response_var,
                    "Victim.age" = shark_attacks_timeline,
                    "attack_count" = shark_attacks_count)
})

lineplot <- c("Victim.age", "attack_count")
barchart <- c("Victim.injury", "Site.category", "Provoked.unprovoked", "Victim.activity", "Injury.severity", "Victim.gender")


# try changing the categories to x axis and time to color
#jerzy prefer's the line and point graphs

renderPlotly({
  if(input$response_var %in% barchart){
    barchart_plot <- ggplot(dataframe_barchart(), aes_string(y = ifelse(input$prop_count == "prop", paste0("proportion_", input$response_var), ifelse(input$prop_count == "count", paste0("count_", input$response_var), "")), fill = input$response_var)) +
      geom_line(aes_string(x = "Year", group = input$response_var, color = input$response_var))+
      geom_point(aes(x = Year))+

      # barchart_plot <- ggplot(dataframe_barchart(), aes_string(y = ifelse(input$prop_count == "prop", paste0("proportion_", input$response_var), ifelse(input$prop_count == "count", paste0("count_", input$response_var), "")), fill = "Year")) +
      # geom_line(aes_string(x = input$response_var, group = "Year", color = "Year"))+
      # geom_point(aes_string(x = input$response_var, group = "Year", color = "Year"))+
        # geom_bar(aes(x = Year), stat = "identity", position = position_dodge(preserve = "single")) +
         labs(x = "Year",
              y = ifelse(input$prop_count == "prop", "Proportion","Count"), 
              title = ifelse(input$response_var == "Victim.injury", "Proportion of Victim Injuries Severity Over Time", ifelse(input$response_var == "Site.category", "Proportion of Shark Attack Site Category Over Time", ifelse(input$response_var == "Provoked.unprovoked", "Proportion of Provoked vs. Unprovoked Shark Attacks Over Time", ifelse(input$response_var == "Victim.activity", "Proportion of Victim Activities Over Time", ifelse(input$response_var == "Injury.severity", "Proportion of Injurity Severity Over Time", ifelse(input$response_var == "Victim.gender", "Proportion of Victim Genders Over Time", "")))))),
              fill = ifelse(input$response_var == "Victim.injury", "Victim Injuries", ifelse(input$response_var == "Site.category", "Site Categories", ifelse(input$response_var == "Provoked.unprovoked", "Provoked?", ifelse(input$response_var == "Victim.activity", "Victim Activities", ifelse(input$response_var == "Injury.severity", "Injurity Severity", ifelse(input$response_var == "Victim.gender", "Victim Gender", ""))))))),
      scale_color_manual(values = c("Fatal" = "#a50f15", "Injured" = "#ef3b2c", "Uninjured" = "#fcbba1", "Island Open Ocean" = "#045a8d", "Estuary/Habor" = "#2b8cbe", "Coastal" = "#74a9cf",  "River" = "#a6bddb", "Other: Fish Farm" = "#d0d1e6","Provoked" = "#8c3a40", "Unprovoked" = "#343d6b", "Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6",
          "Major Lacerations" = "#a50f15", "Minor Lacerations" = "#de2d26", "Punctures" = "#fb6a4a" , "Surface Wound" = "#fc9272", "Abrasion" = "#fcbba1", "Other: Teeth Marks" = "#fee5d9", "Male" = "#a6cee3", "Female" = "#fb9a99"))+

       scale_fill_manual(values = c("Fatal" = "#a50f15", "Injured" = "#ef3b2c", "Uninjured" = "#fcbba1", "Island Open Ocean" = "#045a8d", "Estuary/Habor" = "#2b8cbe", "Coastal" = "#74a9cf",  "River" = "#a6bddb", "Other: Fish Farm" = "#969696","Provoked" = "#8c3a40", "Unprovoked" = "#343d6b", "Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6",
          "Major Lacerations" = "#a50f15", "Minor Lacerations" = "#de2d26", "Punctures" = "#fb6a4a" , "Surface Wound" = "#fc9272", "Abrasion" = "#fcbba1", "Other: Teeth Marks" = "#fee5d9", "Male" = "#a6cee3", "Female" = "#fb9a99")) +
        
        # scale_color_brewer(palette = "Greys") +
        # scale_fill_brewer(palette = "Greys") +
      interactive_theme +
      theme(legend.position = "bottom")
     # Add error bars if selected
  if (input$error_bar & input$prop_count == "prop") {
    barchart_plot <- barchart_plot +
      geom_errorbar(data = dataframe_barchart(), 
                    aes_string(x = "Year", 
                               ymin = paste0("lower_", input$response_var), 
                               ymax = paste0("upper_", input$response_var)), 
                    color = "red", position_dodge(preserve = "single"))# position_dodge(0.9)) #position = "dodge") #position_dodge(preserve = "single")
    
    } 

    
    return(ggplotly(barchart_plot)) 
   
  } else if (input$response_var %in% lineplot) {
    lineplot_plot <- ggplot(data = dataframe_time(),
                            aes_string(x = "Incident.year", y = input$response_var)) +
        geom_line()+
        geom_point() +
        geom_smooth(method = "loess", alpha = 0.2, color = alpha("blue", 0.3)) +
        labs(x = "Date", y = ifelse(input$response_var == "attack_count", "Number of Attacks", "Victim Age")) +
        scale_x_date(date_breaks = "10 years", date_labels = ifelse(input$response_var == "attack_count", "%Y", "%Y"), limits = start.end) +
        #stat_rollapplyr(width = 50, align = "right", linewidth = 1, color = "red") +
        theme(axis.text.x = element_text(angle = 45, vjust = 0.8)) +
      interactive_theme
    
    return(ggplotly(lineplot_plot))
   
  }
  
})


```



Column {.sidebar}
-----------------------------------------------------------------------



```{r}


#widgets for the barcharts

radioButtons("response_var", label = h3("Attack Characteristic:"), 
    choices = list("Number of Attacks" = "attack_count", "Victim Age" = "Victim.age", "Victim Injury" = "Victim.injury", "Site Category" = "Site.category", "Provoked?" = "Provoked.unprovoked", "Victim Activity" = "Victim.activity","Injury Severity" = "Injury.severity", "Victim Gender" = "Victim.gender"), selected = "Victim.injury")

radioButtons("prop_count", label = h3("Proportions or Count:"), 
                   choices = list("Proportion" = "prop", "Count" = "count"))

checkboxInput("error_bar", label = h4("Include Error Bar"))

```


Where are Shark Attacks Occuring? {data-navmenu="Menu"}
=====================================  

```{r}
#loading in the sf data

#removing the spaces in the latitude and longitude data that causes errors
shark_attacks_fixed_latlon <- shark_attacks %>%
  mutate(Latitude = str_replace_all(Latitude, "\\s+", "")) %>%
  mutate(Longitude = str_replace_all(Longitude, "\\s+", "")) %>%
  mutate(Latitude = as.numeric(Latitude)) %>%
  mutate(Longitude = as.numeric(Longitude)) %>%
  filter(is.na(Latitude) == F) %>%
  filter(is.na(Longitude) == F)


# To make the map with points
shark_attacks_sf <- shark_attacks_fixed_latlon %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(crs = 3577)


#crop the australia polygon 
aus_shp <- st_crop(aus_shp, xmin = 113, ymin = -43.9, xmax = 153.9, ymax = -10) #100, ymin = -46, xmax = 160, ymax = -12)   

#aus_shp <- aus_shp$geometry #assigning only the geometry of the states
aus_shp <- st_as_sf(st_transform(aus_shp, crs = 3577))

#filtering out other territories
aus_shp <- aus_shp %>%
  filter(STE_NAME21 != "Other Territories")




```

Sidebar {.sidebar}
-----------------------------------------------------------------------

```{r}
#widgets for the map of australia 

# sliderInput("point_transp", label = h3("Point Transparency"), min = 0, 
#         max = 5, value = 1, step = 0.1)
# 
# sliderInput("point_size", label = "Point size:",
#              min = 0.5, max = 3, value = 2, step = 0.5)

radioButtons("response_var_map", label = h3("Attack Characteristic:"),
    choices = list("Victim Injury" = "Victim.injury", "Site Category" = "Site.category", "Provoked?" = "Provoked.unprovoked", "Victim Activity" = "Victim.activity","Injury Severity" = "Injury.severity", "Victim Gender" = "Victim.gender"), selected = "Victim.injury")

# radioButtons("response_var_map", label = h3("Attack Characteristic:"),
#     choices = list("New South Wales" = "Victim.injury", "Site Category" = "Site.category", "Provoked?" = "Provoked.unprovoked", "Victim Activity" = "Victim.activity","Injury Severity" = "Injury.severity", "Victim Gender" = "Victim.gender"), selected = "Victim.injury")





```

Column 
-----------------------------------------------------------------------

```{r}

aus_shp <- sf::st_cast(aus_shp, "MULTIPOLYGON") #group_or_split = T
shark_attacks_sf <- sf::st_cast(shark_attacks_sf, "MULTIPOINT")

renderPlotly({
  map <- ggplot()+
    geom_sf(data = aus_shp, aes(fill = STE_NAME21)) +
    labs(x = "", y = "", fill = "State") +
    geom_sf_text(data = aus_shp, aes(label = STE_NAME21), color = "black", size = 2, check_overlap = T) +
    geom_sf(data = shark_attacks_sf, aes(fill = .data[[input$response_var_map]], text = paste0("Date: ", paste(Incident.month, Incident.year, sep = "/"), 
                      "<br>Location: ", Location,
                      "<br>Victim Activity:" , Victim.activity,
                      "<br>Injury Type: ", Victim.injury,
                      "<br>Victim Gender: ", Victim.gender,
                      "<br>Site Category: ", Site.category,
                      "<br>Injury Severity: ", Injury.severity)), 
          size = 1.5) +
    scale_color_manual(values = c("Fatal" = "#a50f15", "Injured" = "#ef3b2c", "Uninjured" = "#fcbba1", "Island Open Ocean" = "#045a8d", "Estuary/Habor" = "#2b8cbe", "Coastal" = "#74a9cf",  "River" = "#a6bddb", "Other: Fish Farm" = "#d0d1e6","Provoked" = "#8c3a40", "Unprovoked" = "#343d6b", "Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6",
          "Major Lacerations" = "#a50f15", "Minor Lacerations" = "#de2d26", "Punctures" = "#fb6a4a" , "Surface Wound" = "#fc9272", "Abrasion" = "#fcbba1", "Other: Teeth Marks" = "#fee5d9", "Male" = "#a6cee3", "Female" = "#fb9a99"))+

       scale_fill_manual(values = c("Fatal" = "#a50f15", "Injured" = "#ef3b2c", "Uninjured" = "#fcbba1", "Island Open Ocean" = "#045a8d", "Estuary/Habor" = "#2b8cbe", "Coastal" = "#74a9cf",  "River" = "#a6bddb", "Other: Fish Farm" = "#969696","Provoked" = "#8c3a40", "Unprovoked" = "#343d6b", "Swimming" = "#a6cee3", "Boarding" = "#1f78b4", "Spearfishing" = "#b2df8a", "Snorkeling" = "#33a02c", "Diving" = "#fb9a99", "Boating" = "#e31a1c", "Standing in Water" = "#fdbf6f", "Fishing" = "#ff7f00", "Other" = "#cab2d6",
          "Major Lacerations" = "#a50f15", "Minor Lacerations" = "#de2d26", "Punctures" = "#fb6a4a" , "Surface Wound" = "#fc9272", "Abrasion" = "#fcbba1", "Other: Teeth Marks" = "#fee5d9", "Male" = "#a6cee3", "Female" = "#fb9a99")) +
      interactive_theme
    
  
  return(ggplotly(map))

})


```

