---
title: "REBECCA Practice"
output: flexdashboard::flex_dashboard
runtime: shiny
---

## Actually running and deploying the app
```{r}
#running the demo app
#runApp()

```

# Loading in the Data and Packages
```{r global, include=FALSE, warning=FALSE, message=FALSE}
shark_attacks <- read.csv("Shark Dataset.csv")

# load packages in 'global' chunk so they can be shared by all users of the dashboard
library(httr2)
library(purrr)
library(tibble)
library(lubridate)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyr)
library(ggseas) #to use stat_rollapplyr

```

- first reclassify dates (1790-1850, 1850-1900, 1900-1950, 
- most common location, victim activity, site.category, distance to shore over time, proportion barchart
- 


# cleaning the data

```{r}

#adding in means
shark_attacks <- shark_attacks %>%
  filter(is.na(UIN) != T) %>% #removing the empty rows
  mutate(Victim.injury = case_when(Victim.injury == "fatal" ~ "Fatal",
                                   Victim.injury == "injured" ~ "Injured",
                                   Victim.injury == "Injured" ~ "Injured",
                                   Victim.injury == "uninjured" ~ "Uninjured")) %>%
  mutate(Provoked.unprovoked = case_when(Provoked.unprovoked == "provoked" ~ "Provoked",
                                   Provoked.unprovoked == "unprovoked" ~ "Unprovoked")) %>%
  mutate(Victim.gender = case_when(Victim.gender == "female" ~ "Female",
                                   Victim.gender == "male" ~ "Male")) %>%
  mutate(Injury.severity = case_when(Injury.severity == "abrasion" ~ "Abrasion", Injury.severity == "major lacerations" ~ "Major Lacerations", Injury.severity == "minor lacerations" ~ "Minor Lacerations", Injury.severity == "other: teeth marks" ~ "Other: Teeth Marks", Injury.severity == "punctures" ~ "Punctures", Injury.severity == "surface wound" ~ "Surface Wound")) %>%
  mutate(Site.category = case_when(Site.category == "coastal" ~ "Coastal",
                                   Site.category == "Coastal" ~ "Coastal",
                                   Site.category == "estuary/harbour" ~ "Estuary/Harbour",
                                   Site.category == "island open ocean" ~ "Island Open Ocean", 
                                   Site.category == "ocean/pelagic" ~ "Ocean/Pelagic",
                                   Site.category == "other: fish farm" ~ "Other: Fish Farm",
                                   Site.category == "river" ~ "River")) %>%
  filter(Incident.year != "1791") %>%
  mutate(Victim.activity = case_when(Victim.activity == "boarding" ~ "Boarding",
                                   Victim.activity == "diving" ~ "Diving",
                                   Victim.activity == "fishing" ~ "Fishing",
                                   Victim.activity == "motorised boating" ~ "Motorised Boating",
                                   Victim.activity == "other: standing in water" ~ "Standing in Water",
                                   Victim.activity == "other:floating" ~ "Floating",
                                   Victim.activity == "snorkeling" ~ "Snorkeling", 
                                   Victim.activity == "spearfishing" ~ "Spearfishing",
                                   Victim.activity == "swimming" ~ "Swimming",
                                   Victim.activity == "unmotorised boating" ~ "Unmotorised Boating")) %>%
  filter(Incident.year != "1791") %>%
  group_by(Incident.year) %>%
   #creating mean columns
  mutate(Shark.length.m.mean = mean(Shark.length.m)) %>%
  mutate(No.sharks.mean = mean(No.sharks)) %>%
  mutate(Victim.age.mean = mean(Victim.age)) %>%
  #creating date groupings
  mutate(Year = case_when((Incident.year > 1800 & Incident.year <= 1850) ~ "1800-1850", 
                          (Incident.year > 1850 & Incident.year <= 1900) ~ "1850-1900",
                          (Incident.year > 1900 & Incident.year <= 1950) ~ "1900-1950",
                          (Incident.year > 1950 & Incident.year <= 2000) ~ "1950-2000",
                          (Incident.year > 2000 & Incident.year <= 2022) ~ "2000-2022")) 

#counting the number of sharks every 50 years
shark_attacks <- shark_attacks %>%
  group_by(Year) %>%
  mutate(Number_Attacks = n())
```



# computing proportions individually

```{r}
#individually

#Victim.injury
shark_attacks_proportions_Victim.injury <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.injury != "") %>%
  group_by(Year, Victim.injury) %>%  #  group by the year variable
  summarise(count_Victim.injury = n()) %>% 
  mutate(total_Victim.injury = sum(count_Victim.injury),  #  compute total number of observations
         proportion_Victim.injury = count_Victim.injury / total_Victim.injury,  #  compute proportions
         std_error_Victim.injury = sqrt(proportion_Victim.injury * (1-proportion_Victim.injury) / total_Victim.injury),  #  compute standard error of each proportion
         lower_Victim.injury = proportion_Victim.injury - 1.96 * std_error_Victim.injury,  #  compute CI lower bound
         upper_Victim.injury = proportion_Victim.injury + 1.96 * std_error_Victim.injury) 
  

#Site.category
shark_attacks_proportions_Site.category <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Site.category != "") %>%
  group_by(Year, Site.category) %>%  #  group by the year variable
  summarise(count_Site.category = n()) %>%  
  mutate(total_Site.category = sum(count_Site.category),  
         proportion_Site.category = count_Site.category / total_Site.category,  
         std_error_Site.category = sqrt(proportion_Site.category * (1-proportion_Site.category) / total_Site.category),  
         lower_Site.category = proportion_Site.category - 1.96 * std_error_Site.category,  
         upper_Site.category = proportion_Site.category + 1.96 * std_error_Site.category)
  

#Provoked.unprovoked
shark_attacks_proportions_Provoked.unprovoked <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Provoked.unprovoked != "") %>%
  group_by(Year, Provoked.unprovoked) %>%  #  group by the year variable
  summarise(count_Provoked.unprovoked = n()) %>%  
  mutate(total_Provoked.unprovoked = sum(count_Provoked.unprovoked),  
         proportion_Provoked.unprovoked = count_Provoked.unprovoked / total_Provoked.unprovoked, 
         std_error_Provoked.unprovoked = sqrt(proportion_Provoked.unprovoked * (1-proportion_Provoked.unprovoked) / total_Provoked.unprovoked),  
         lower_Provoked.unprovoked = proportion_Provoked.unprovoked - 1.96 * std_error_Provoked.unprovoked,  
         upper_Provoked.unprovoked = proportion_Provoked.unprovoked + 1.96 * std_error_Provoked.unprovoked)
  

#Victim.activity
shark_attacks_proportions_Victim.activity <- shark_attacks %>%  #  Start with the cereal data.frame
  filter(Victim.activity != "") %>%
  group_by(Year, Victim.activity) %>%  #  group by the year variable
  summarise(count_Victim.activity = n()) %>%  
  mutate(total_Victim.activity = sum(count_Victim.activity),  
         proportion_Victim.activity = count_Victim.activity / total_Victim.activity,  
         std_error_Victim.activity = sqrt(proportion_Victim.activity * (1-proportion_Victim.activity) / total_Victim.activity),  
         lower_Victim.activity = proportion_Victim.activity - 1.96 * std_error_Victim.activity,  
         upper_Victim.activity = proportion_Victim.activity + 1.96 * std_error_Victim.activity)


#Victim.gender 
shark_attacks_proportions_Victim.gender <- shark_attacks %>% #  Start with the data.frame
  filter(Victim.gender != "") %>%
  group_by(Year, Victim.gender) %>%  #  group by the year variable
  summarise(count_Victim.gender = n()) %>%    
  mutate(total_Victim.gender = sum(count_Victim.gender),  
         proportion_Victim.gender = count_Victim.gender / total_Victim.gender,  
         std_error_Victim.gender = sqrt(proportion_Victim.gender * (1-proportion_Victim.gender) / total_Victim.gender), 
         lower_Victim.gender = proportion_Victim.gender - 1.96 * std_error_Victim.gender,  
         upper_Victim.gender = proportion_Victim.gender + 1.96 * std_error_Victim.gender)
  
  
#Injury.severity
shark_attacks_proportions_Injury.severity <- shark_attacks %>%  #Start with the cereal data.frame
  filter(Injury.severity != "") %>%
  group_by(Year, Injury.severity) %>%  #  group by the year variable
  summarise(count_Injury.severity = n()) %>%  
  mutate(total_Injury.severity = sum(count_Injury.severity),  
         proportion_Injury.severity = count_Injury.severity / total_Injury.severity,  
         std_error_Injury.severity = sqrt(proportion_Injury.severity * (1-proportion_Injury.severity) / total_Injury.severity), 
         lower_Injury.severity = proportion_Injury.severity - 1.96 * std_error_Injury.severity,  
         upper_Injury.severity = proportion_Injury.severity + 1.96 * std_error_Injury.severity) 


shark_attacks_proportions_Injury.severity <- shark_attacks %>%  #Start with the cereal data.frame
  filter(Injury.severity != "") %>%
  group_by(Year, Injury.severity) %>%  #  group by the year variable
  summarise(count_Injury.severity = n(),
            total_Injury.severity = sum(count_Injury.severity),  
            proportion_Injury.severity = count_Injury.severity / total_Injury.severity,
            std_error_Injury.severity = sqrt(proportion_Injury.severity * (1-proportion_Injury.severity) / total_Injury.severity), 
            lower_Injury.severity = proportion_Injury.severity - 1.96 * std_error_Injury.severity,
            upper_Injury.severity = proportion_Injury.severity + 1.96 * std_error_Injury.severity) 

```



Introduction: You Won a Trip to Australia!
=====================================  


How Have Shark Attacks Changed over Time?
=====================================  

### Shark Attacks over Time

```{r}

dataframe <- reactive({switch(input$response_var,
                    "Victim.injury" = shark_attacks_proportions_Victim.injury,
                    "Site.category" = shark_attacks_proportions_Site.category,
                    "Provoked.unprovoked" = shark_attacks_proportions_Provoked.unprovoked,
                    "Victim.activity" = shark_attacks_proportions_Victim.activity,
                    "Injury.severity" = shark_attacks_proportions_Injury.severity,
                     "Victim.gender" = shark_attacks_proportions_Victim.gender)
})


renderPlot({
  ggplot(dataframe(), aes_string(y = paste0("proportion_", input$response_var), fill = input$response_var)) +
         geom_bar(aes(x = Year), stat = "identity", position = "dodge") + 
         if(input$error_bar == T){geom_errorbar(aes(ymin = lower_Victim.injury, ymax = upper_Victim.injury), color = "red", position = "dodge")} + 
         labs(x = "HS Graduation Rate",
              title = "Education vs Illiteracy in US States in 1977")
})
```



Column {.sidebar}
-----------------------------------------------------------------------


```{r}


# sliderInput("point_size", label = "Point size:",
#             min = 0.5, max = 3, value = 2, step = 0.5)

checkboxGroupInput("response_var", label = h3("Response Variable:"), 
    choices = list("Victim Injury" = "Victim.injury", "Site Category" = "Site.category", "Provoked?" = "Provoked.unprovoked", "Victim Activity" = "Victim.activity","Injury Severity" = "Injury.severity", "Victim Gender" = "Victim.gender"), selected = 1)

#"Victim Age" = "Victim.age", "Shark Length" = "Shark.length.m", "Number of Sharks During the Attacks" = "No.sharks", 
  
    
    #("Victim Age" = 1, "Shark Length" = 2, "Number of Sharks During the Attacks" = 3, "Victim Injury" = 4, "Site Category" = 5, "Site Category" = 6, "Provoked?" = 7, "Victim Activity" = 8,"Injury Severity" = 9, "Victim Gender" = 10), selected =1)
    

checkboxInput("error_bar", label = "Include Error Bar")
```




